---
title: "Global Coral Microbiome Project: Statistical Analysis and Modelling Workflow 2024"
authors: "Colin Howe, Sofia Roitman"
start date: "01/17/2024"
current date: "03/29/2024"
output: html_document
---

# First, Set working Directory load library and Import Data into Phyloseq
```{r}
# Command for setting your work directory
knitr::opts_knit$set("C:/Users/Colin Howe/Desktop/PSU Doctoral Program/GCMP_Working/gcmp_analysis_2024/Analysis/Neutral_Model_Analysis/a1_neutrality_by_compartment_all")
                 
# Print your working directory

```

## Load required libraries. 
```{r, message=FALSE, warning=FALSE} 
library(BiocManager)
library(phyloseq)
library(tidyverse)
library(qiime2R)
library(DESeq2)
library(ComplexHeatmap)
library(ade4)
library(vegan)
library(ggplot2)
library(microbiome)
library(dendextend)
library(RVAideMemoire)
library(ape)
library(RColorBrewer)
library(ggsignif)
library(ANCOMBC)
library(biomformat)
```


# Import OTU Table Taxonomy and Tree from Qiime2 and import Metadata
```{r,message=FALSE, warning=FALSE}
#Import from .qza file into a phyloseq object
asv <- qza_to_phyloseq(features = "./merged_table.qza")

#### Import Metadata read.table
metadata <- read.table(file = "./gcmp_complete_mapping2025.txt",header=T,comment.char="", row.names=1, sep="\t")

### Import Tree file from biom output tree.nwk
tree <- read_tree("./tree.nwk")

### Import taxonomy from biom output as .tsv format using read.table
taxonomy <- read.table(file = "./taxonomy.tsv", sep = "\t", header = T ,row.names = 1)

# clean the taxonomy
##code referenced from Yan Hui: email me@yanh.org github: yanhui09
tax <- taxonomy %>%
  select(Taxon) %>% 
  separate(Taxon, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), "; ")

tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "k__",""),
                        Phylum = str_replace(tax[,2], "p__",""),
                        Class = str_replace(tax[,3], "c__",""),
                        Order = str_replace(tax[,4], "o__",""),
                        Family = str_replace(tax[,5], "f__",""),
                        Genus = str_replace(tax[,6], "g__",""),
                        Species = str_replace(tax[,7], "s__",""),
                        stringsAsFactors = FALSE)

tax.clean[is.na(tax.clean)] <- ""
tax.clean[tax.clean=="__"] <- ""

for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,7] != ""){
    tax.clean$Species[i] <- paste(tax.clean$Genus[i], tax.clean$Species[i], sep = " ")
  } else if (tax.clean[i,2] == ""){
    kingdom <- paste("Unclassified", tax.clean[i,1], sep = " ")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Unclassified", tax.clean[i,2], sep = " ")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Unclassified", tax.clean[i,3], sep = " ")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Unclassified", tax.clean[i,4], sep = " ")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Unclassified", tax.clean[i,5], sep = " ")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Unclassified ",tax.clean$Genus[i], sep = " ")
  }
}

# create matrix format for OTU and taxonomy table
OTU <- otu_table(as.matrix(asv), taxa_are_rows = TRUE)
tax1 = tax_table(as.matrix(tax.clean))

# Set metadata
SAMPLE <- sample_data(metadata)

# Create Working phyloseq object
main_phylo <- phyloseq(OTU,tax1,SAMPLE,tree)

# visualize tax table
table(tax_table(main_phylo)[,"Kingdom"])

# remove unknown bacteria or unassigned
phylo_noking <-main_phylo %>%
  phyloseq::subset_taxa(!Kingdom %in% c("Unassigned","Unclassified d__Bacteria","d__Eukaryota")) %>%
  phyloseq::subset_taxa(!Phylum %in% c("Unclassified Unassigned","Unclassified d__Archaea","Unclassified d__Bacteria","Unclassified d__Eukaryota")) %>%
  phyloseq::subset_taxa(!Family %in% c("Mitochondria","Chloroplast"))

# verify present taxa in table
table(tax_table(phylo_noking)[,"Kingdom"])

# Remove any samples with less than 1 read
phylo = prune_samples(sample_sums(phylo_noking)>1, phylo_noking)
phylo

# Subset only corals from database
# Subset only corals from database
subject <-subset_samples(phylo, outgroup == "n" & tissue_compartment == "T")
subject

print(paste("Generating Rarefied Coral dataset..."))
# rarefy to even depth
rarefied = rarefy_even_depth(subject, rngseed=111, sample.size=1000, replace=F, trimOTUs = TRUE)
rarefied

###Agglomerate taxa to family 
glom <- tax_glom(rarefied, taxrank = 'Family', NArm = TRUE)
glom

```


#Bruns et al. 2016 Neutral Model functions
```{r}
################define sncm.fit############
sncm.fit <- function(spp, pool=NULL, stats=TRUE, taxon=NULL){
	require(minpack.lm)
	require(Hmisc)
	require(stats4)
	
	options(warn=-1)

	#Calculate the number of individuals per community
	N <- mean(apply(spp, 1, sum))
	
	#Calculate the average relative abundance of each taxa across communities
	if(is.null(pool)){
		p.m <- apply(spp, 2, mean)
		p.m <- p.m[p.m != 0]
		p <- p.m/N
	} else {
		p.m <- apply(pool, 2, mean)
		p.m <- p.m[p.m != 0]
		p <- p.m/N
	}

#Calculate the occurrence frequency of each taxa across communities
spp.bi <- 1*(spp>0)
freq <- apply(spp.bi, 2, mean)
freq <- freq[freq != 0]
#Combine
C <- merge(p, freq, by=0)
C <- C[order(C[,2]),]
C <- as.data.frame(C)
C.0 <- C[!(apply(C, 1, function(y) any(y == 0))),] #Removes rows with any zero (absent in either source pool or local communities)
p <- C.0[,2]
freq <- C.0[,3]
names(p) <- C.0[,1]
names(freq) <- C.0[,1]

#Calculate the limit of detection
d = 1/N

##Fit model parameter m (or Nm) using Non-linear least squares (NLS)
m.fit <- nlsLM(freq ~ pbeta(d, N*m*p, N*m*(1-p), lower.tail=FALSE), start=list(m=0.1))
m.ci <- confint(m.fit, 'm', level=0.95)
	
##Fit neutral model parameter m (or Nm) using Maximum likelihood estimation (MLE)
sncm.LL <- function(m, sigma){
	R = freq - pbeta(d, N*m*p, N*m*(1-p), lower.tail=FALSE)
	R = dnorm(R, 0, sigma)
	-sum(log(R))
}
	
m.mle <- mle(sncm.LL, start=list(m=0.1, sigma=0.1), nobs=length(p),"Nelder-Mead")

##Calculate Akaike's Information Criterion (AIC)
aic.fit <- AIC(m.mle, k=2)
bic.fit <- BIC(m.mle)

##Calculate goodness-of-fit (R-squared and Root Mean Squared Error)
freq.pred <- pbeta(d, N*coef(m.fit)*p, N*coef(m.fit)*(1-p), lower.tail=FALSE)
Rsqr <- 1 - (sum((freq - freq.pred)^2))/(sum((freq - mean(freq))^2))
RMSE <- sqrt(sum((freq-freq.pred)^2)/(length(freq)-1))
	
pred.ci <- binconf(freq.pred*nrow(spp), nrow(spp), alpha=0.01, method="wilson", return.df=TRUE)
	
##Calculate AIC for binomial model
bino.LL <- function(mu, sigma){
R = freq - pbinom(d, N, p, lower.tail=FALSE)
R = dnorm(R, mu, sigma)
-sum(log(R))
	}
bino.mle <- mle(bino.LL, start=list(mu=0, sigma=0.1), nobs=length(p))
	
aic.bino <- AIC(bino.mle, k=2)
bic.bino <- BIC(bino.mle)
	
##Goodness of fit for binomial model
bino.pred <- pbinom(d, N, p, lower.tail=FALSE)
Rsqr.bino <- 1 - (sum((freq - bino.pred)^2))/(sum((freq - mean(freq))^2))
RMSE.bino <- sqrt(sum((freq - bino.pred)^2)/(length(freq) - 1))

bino.pred.ci <- binconf(bino.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)
	
##Calculate AIC for Poisson model
pois.LL <- function(mu, sigma){
	R = freq - ppois(d, N*p, lower.tail=FALSE)
	R = dnorm(R, mu, sigma)
	-sum(log(R))
}
pois.mle <- mle(pois.LL, start=list(mu=0, sigma=0.1), nobs=length(p))
	
aic.pois <- AIC(pois.mle, k=2)
bic.pois <- BIC(pois.mle)
	
##Goodness of fit for Poisson model
pois.pred <- ppois(d, N*p, lower.tail=FALSE)
Rsqr.pois <- 1 - (sum((freq - pois.pred)^2))/(sum((freq - mean(freq))^2))
RMSE.pois <- sqrt(sum((freq - pois.pred)^2)/(length(freq) - 1))

pois.pred.ci <- binconf(pois.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)

##Results
	if(stats==TRUE){
		fitstats <- data.frame(m=numeric(), m.ci=numeric(), m.mle=numeric(), maxLL=numeric(), binoLL=numeric(), poisLL=numeric(), Rsqr=numeric(), Rsqr.bino=numeric(), Rsqr.pois=numeric(), RMSE=numeric(), RMSE.bino=numeric(), RMSE.pois=numeric(), AIC=numeric(), BIC=numeric(), AIC.bino=numeric(), BIC.bino=numeric(), AIC.pois=numeric(), BIC.pois=numeric(), N=numeric(), Samples=numeric(), Richness=numeric(), Detect=numeric())
		fitstats[1,] <- c(coef(m.fit), coef(m.fit)-m.ci[1], m.mle@coef['m'], m.mle@details$value, bino.mle@details$value, pois.mle@details$value, Rsqr, Rsqr.bino, Rsqr.pois, RMSE, RMSE.bino, RMSE.pois, aic.fit, bic.fit, aic.bino, bic.bino, aic.pois, bic.pois, N, nrow(spp), length(p), d)
		return(fitstats)
	} else {
	A <- cbind(p, freq, freq.pred, pred.ci[,2:3], bino.pred, bino.pred.ci[,2:3])
	A <- as.data.frame(A)
		colnames(A) <- c('p_abundance', 'freq', 'freq.pred', 'pred.lwr', 'pred.upr', 'bino.pred', 'bino.lwr', 'bino.upr')
		if(is.null(taxon)){
			B <- A[order(A[,1]),]
		} else {
			B <- merge(A, taxon, by=0, all=TRUE)
			row.names(B) <- B[,1]
			B <- B[,-1]
			B <- B[order(B[,1]),]
		}
		return(B)
	}
}
##Identify above, below and neutral OTUs 

addf = function(freq, pred.upr, pred.lwr, padj) {
  if(freq > pred.upr){
    model = "above"
  }
  else if(freq < pred.lwr){
    model = "below"
  }
  else if(pred.upr >= freq & freq >= pred.lwr){
    model = "neutral"
  }
  return(model)
}
```

## Neutral Model test
```{r}
###coral communities
set.seed(111)
coral <- prune_taxa(taxa_sums(glom) > 0, glom)
coral
otu_coral = as.data.frame(t(otu_table(coral)))

## Neutral fit
mod_w_coral = sncm.fit(otu_coral)
mod1_w_coral = sncm.fit(otu_coral,stats = F)

min(sample_sums(glom))

mod1_w_coral %>% mutate(ci.mean = (pred.upr-pred.lwr)/2,
                        SE = (ci.mean/(1.96)),
                        zstat = (freq/SE),
                        p_value = 2*pnorm(-abs(zstat)),
                        #p_value = (exp(-0.717*zstat-0.416*zstat^2)),
                        padj = p.adjust(p_value, method = "bonferroni", n=length(mod1_w_coral$freq))) %>% as.data.frame() %>%
                        rownames_to_column("id")-> neutral_results_padj
                        #padj = (exp(-0.717*zstat-0.416*zstat^2))*length(mod1_w_coral$model)) -> neutral_results_padj

neutral_results_padj$model = mapply(addf, neutral_results_padj$freq, neutral_results_padj$pred.upr, neutral_results_padj$pred.lwr, neutral_results_padj$padj)  

length(which(neutral_results_padj$model == "above"))
length(which(neutral_results_padj$model == "below"))
length(which(neutral_results_padj$model == "neutral"))


neutral_results_padj$model <- factor(neutral_results_padj$model, 
                             levels = c("above", "neutral", "below"))

################# seawater ####################
set.seed(111)
###seawater communities
#seawater <- subset_samples(physeq2, SampleType == "seawater")
seawater <- prune_taxa(taxa_sums(glom_col) >0, glom_col)
seawater

otu_seawater = as.data.frame(t(otu_table(seawater)))

mod_w_seawater = sncm.fit(otu_seawater)
mod1_w_seawater = sncm.fit(otu_seawater,stats = F)


mod1_w_seawater$model = mapply(addf, mod1_w_seawater$freq, mod1_w_seawater$pred.upr, mod1_w_seawater$pred.lwr)  
length(which(mod1_w_seawater$model == "above"))
length(which(mod1_w_seawater$model == "below"))
length(which(mod1_w_seawater$model == "neutral"))

mod1_w_seawater$model <- factor(mod1_w_seawater$model, 
                                levels = c("above", "neutral", "below"))

## Assign Taxonomy to table of significant microbes
#env_taxa = cbind(as(mod1_w_seawater, "data.frame"), as(tax_table(glom_col12)[rownames(mod1_w_seawater), ], "matrix"))

#write.table(env_taxa,"Colombia E12 Samples Neutral Taxonomy table.tsv",row.names = TRUE)


```

## Plot Neutral Model
```{r}
#plot neuModel
p_neuM_coral=ggplot(neutral_results_padj, aes(x=log10(p_abundance), y=freq, color = model))+
  geom_point(shape=19, alpha=0.7, size=2.5)+
  geom_line(aes(x=log10(p_abundance), y=freq.pred),
            color="blue",linetype="solid",linewidth=1)+
  geom_line(aes(x=log10(p_abundance), y=pred.lwr),
            color="blue",linetype="dashed",linewidth=1)+
  geom_line(aes(x=log10(p_abundance), y=pred.upr),
            color="blue",linetype="dashed",linewidth=1)+
    scale_colour_manual(name="OTUs",
                      values=c("#E41A1C", "#999999", "#FF7F00","purple"),
                      breaks=c("above", "neutral", "below","neutral_significant"),
                      labels=c("Over-represented", "Neutrally-distributed", "Under-represented","Significant_low_weight"))+
  xlab("Log10 (Mean Relative abundance)") + ylab("Occurance frequency")+
  theme_bw() +
  theme(axis.title.x = element_text(face="bold", size=15),
        axis.title.y = element_text(face="bold", size=15),
        axis.text.x = element_text(colour = "black", size=12),
        axis.text.y = element_text(colour = "black", size=12), 
        legend.title = element_text(size=11.5, face = "bold"),
        legend.text = element_text(size=11.5),
        legend.position="none")

#expedition <-"E5"
biosample <-"Skeleton"
r <- round(mod_w_coral$Rsqr,3)
m_var <- round(mod_w_coral$m,3)
lb1 <- paste0("R_squared",":",r)
lb2 <- paste0("migration_rate (m):",m_var)
lb3 <- paste0(biosample,"_Neutral_Model")#expedition,"_",

p_neuM_coral1 = p_neuM_coral + annotate("text", x = -3.5, y=0.7, size=6.5, label=lb1, 
                                        fontface="bold", color="black",
                                        parse=TRUE) +
  annotate("text", x = -3.5, y=0.6, size=6.5, label=lb2, 
           fontface="bold", color="black",
           parse=TRUE) +
  annotate("text", x = -3.5, y=0.9, size=6.5, label=lb3, 
           fontface="bold", color="black",
           parse=FALSE)

#ggsave("Tissue_sig_neutral_model_plot.jpg", p_neuM_coral1, width = 20, height = 10, dpi= 600)

############ Seawater PLot #############
#plot neuModel
p_neuM_seawater=ggplot(mod1_w_seawater, aes(x=log10(p), y=freq, color = model))+
  geom_point(shape=19, alpha=0.7, size=2.5)+
  geom_line(aes(x=log10(p), y=freq.pred),
            color="blue",linetype="solid",size=1)+
  geom_line(aes(x=log10(p), y=pred.lwr),
            color="blue",linetype="dashed",size=1)+
  geom_line(aes(x=log10(p), y=pred.upr),
            color="blue",linetype="dashed",size=1)+
  scale_colour_manual(name="OTUs",
                      values=c("#E41A1C", "#999999", "#FF7F00"),
                      breaks=c("above", "neutral", "below"),
                      labels=c("Over-represented", "Neutrally-distributed", "Under-represented"))+
  xlab("Log10 (Mean Relative abundance)") + ylab("Occurance frequency")+
  theme_bw() +
  theme(axis.title.x = element_text(face="bold", size=15),
        axis.title.y = element_text(face="bold", size=15),
        axis.text.x = element_text(colour = "black", size=12),
        axis.text.y = element_text(colour = "black", size=12), 
        legend.title = element_text(size=11.5, face = "bold"),
        legend.text = element_text(size=11.5),
        legend.position="none")#http://127.0.0.1:24155/graphics/plot_zoom_png?width=1762&height=934

ev1 <- mod_w_seawater$Rsqr
ev2 <- mod_w_seawater$m
ev3 <- expression(paste("Colombia E2 & E12 \n Coral Samples"))

p_neuM_seawater1 = p_neuM_seawater + annotate("text", x = -3.5, y=0.7, size=6.5, label=ev1, 
                                              fontface="bold", color="black",
                                              parse=TRUE) +
annotate("text", x = -3.5, y=0.6, size=6.5, label=ev2, 
           fontface="bold", color="black",
           parse=TRUE) +
annotate("text", x = -3.5, y=0.9, size=5.5, label=ev3, 
           fontface="bold", color="black",
           parse=TRUE)

getwd()

mp_ncm = ggdraw() +
  draw_plot(p_neuM_coral1, x = 0.01, y = 0.1, width = 0.49, height = 0.8) + 
  draw_plot(p_neuM_seawater1, x = 0.51, y = 0.1, width = 0.49, height = 0.8) 
mp_ncm
ggsave(" Caribbean_Mucus Neutral Model Plot.jpg", p_neuM_coral1, width = 20, height = 10, dpi= 600)


```

## Psudo Tables 
```{r}
Tissue_glom_table <-read.table("./Tissue/T_glom_table.tsv",header=T, sep = "\t", check.names=FALSE)
Tissue_glom_tax <-read.table("./Tissue/T_glom_taxonomy.tsv",header=T,sep = "\t", check.names=FALSE)

Tissue_above_map <-read.table("./Tissue/Tissue_above_metadata.txt",header=T,sep = "\t", check.names=FALSE)
Tissue_below_map <-read.table("./Tissue/Tissue_below_metadata.txt",header=T,sep = "\t", check.names=FALSE)
Tissue_neutral_map <-read.table("./Tissue/Tissue_neutral_metadata.txt",header=T,sep = "\t", check.names=FALSE)

rbind(Tissue_above_map,Tissue_below_map,Tissue_neutral_map) -> test_map

## Subset feature table
#phyloseq::otu_table(glom)%>%
#  as.data.frame()%>%
#  rownames_to_column("id") -> Tissue_glom_table

## Subet taxonomy table
#phyloseq::tax_table(glom)%>%
#  as.data.frame() %>%
#  rownames_to_column("id") -> Tissue_glom_tax

## Subet Mapping file
#phyloseq::sample_data(glom) %>% as.data.frame() -> Tissue_glom_mapping


biosample="Tissue"
# Subset Taxonomy and join with neutral model results
neutral = c("above","below")
for (b in neutral){
## Inner_join will connect  ASV id with taxonomy using id column
inner_join(neutral_results_padj,Tissue_glom_tax, by = "id") -> non_neutral_tax

## Subset ASV from each model and signifcance 
sig_neutral_tax = select(filter(non_neutral_tax, model == b & padj <=0.05),id,p_abundance,freq,padj,Phylum,Class,Order,Family,model)

#subset otu table based on neutral results and rename column names
neutral_table <-subset(Tissue_glom_table, id  %in%  sig_neutral_tax$id)
l = length(neutral_table)
colnames(neutral_table)[2:l] <- paste(colnames(neutral_table)[2:l], b, sep = "_")

## Create new columns for the metadata data
T_glom_mapping %>% mutate(psuedo_sample_name = paste(colnames(neutral_table)[2:l]),
                         compartment_neutraility = paste(biosample,b, sep = '_'),
                         compartment = paste (biosample),
                         neutraility = paste(b)) %>% 
                        select(psuedo_sample_name,sample_name_backup, host_scientific_name,expedition_number,political_area,ocean_area,compartment_neutraility,compartment,neutraility) %>% 
                        as.data.frame() -> glom_metadata
## Subset significant results with taxonomy to retain only significant taxonomy
#subset_tax <- subset(Tissue_glom_tax, id %in% sig_neutral_tax$id)# %>% select(-c("Genus","Species"))

psuedo_table_name <- paste0(biosample,"_",b,"_table.tsv")
  (write.table(neutral_table, file =psuedo_table_name,sep="\t", row.names = FALSE))
  
psuedo_map_name <- paste0(biosample,"_",b,"_metadata.csv") ## remember to combine above, below and neutral in excel
  (write.table(glom_metadata, file =psuedo_map_name,sep=",", row.names = FALSE))
  
# psuedo_tax_name <- paste0(biosample,"_",b,"_taxonomy.tsv")
#  (write.table(subset_tax, file =psuedo_tax_name,sep="\t", row.names = FALSE))
}

################## Subset ASV for the neutral taxa ####################
sig_neutral_ref = select(filter(non_neutral_tax, model == "neutral" & padj >=0.05),id,p_abundance,freq,padj,Phylum,Class,Order,Family,model)

#subset otu table based on neutral results and rename column names
neutral_table_test <-subset(Tissue_glom_table, id  %in%  sig_neutral_ref$id)
l = length(neutral_table_test)
colnames(neutral_table_test)[2:l] <- paste(colnames(neutral_table_test)[2:l], "neutral", sep = "_")


### Write neutral tables and mapping file
neutral_table_name <- paste0(biosample,"_neutral_table.tsv")
  (write.table(neutral_table_test, file =neutral_table_name,sep="\t", row.names = FALSE))
  
#neutral_map_name <- paste0(biosample,"_neutral_metadata.csv") ## remember to combine above, below and neutral in excel
#  (write.table(neutral_map_name, file =psuedo_map_name,sep=",", row.names = FALSE))

## Merging above and Below tables for phyloseq object

above_table <- read.table("./Tissue_above_table.tsv",header = TRUE,sep="\t", check.names=FALSE)
below_table <- read.table("./Tissue_below_table.tsv",header = TRUE,sep="\t", check.names=FALSE)
neutral_table <-read.table("./Tissue_neutral_table.tsv",header = TRUE,sep="\t", check.names=FALSE)

### Merge above and below tables
merge(above_table, below_table, by=0,all=T) -> merge_table

merge_table %>% mutate(id = coalesce(id.x,id.y)) %>% 
    relocate(id) %>%  
  select(!c(id.x,id.y,Row.names))-> merge_test
merge_test[is.na(merge_test)] <-0



## Merge with above and below tables
merge(merge_test,neutral_table_test,by=0,all=T) -> full_table
full_table %>% mutate(id = coalesce(id.x,id.y)) %>% 
    relocate(id) %>%  
  select(!c(id.x,id.y,Row.names))-> full_table
full_table[is.na(full_table)] <-0

#write.table(full_table,file = "merged_table_test.txt", sep="\t",row.names = FALSE)
#test_biom <-make_biom(merge_test) ## unknown biom format makes this difficult to import into qiime2
#write_biom(test_biom,biom_file = "tissue_above_biom")

l = length(neutral_table_test)
T_glom_mapping %>% mutate(psuedo_sample_name = paste(colnames(neutral_table_test)[2:l]),
                         compartment_neutraility = paste("Tissue", "neutral", sep = '_'),
                         compartment = paste ("Tissue"),
                         neutraility = paste("neutral")) %>% 
                        select(psuedo_sample_name,sample_name_backup, host_scientific_name,expedition_number,political_area,ocean_area,compartment_neutraility,compartment,neutraility) %>% 
                        as.data.frame() -> glom_metadata

## Subset taxonomy file 
taxonomy2 <- read.table(file = "gcmp_qiime_input/taxonomy.tsv", sep = "\t", header = T)
colnames(taxonomy2)[1] <-"id"
neutral_tax <-subset(taxonomy2, id  %in%  full_table$id)

## Print output files
print(paste("Writing Combined Psudo_table across compartments"))
psudo_table_name <- paste0(biosample,"_combined_psudo_table.tsv")
write.table(merge_tab, file=psudo_table_name, sep="\t",row.names = FALSE)



```

## Neutral Phyloseq Object      
```{r}
## ancombc analysis
library(ANCOMBC)

#Mucus
mucus_table <-read.table("./Mucus/mucus_combined_psudo_table.tsv",header=T,sep="\t",check.names=FALSE)
mucus_tax <-read.table("./Mucus/Mucus_combined_psudo_tax.tsv", header=T,sep="\t",check.names=FALSE)
mucus_mapping <-read.table("./Mucus/Mucus_combined_mapping.csv", header=T,sep=",", check.names=FALSE)

#Tissue
tissue_table <-read.table("./Tissue/Tissue_combined_psudo_table.tsv",row.names=1,header=T,sep="\t",  check.names=FALSE)
tissue_tax <-read.table("./Tissue/Tissue_combined_psudo_tax.tsv",header=T,sep="\t", check.names=FALSE)
tissue_mapping <-read.table("./Tissue/Tissue_combined_mapping.csv",header=T,sep=",",  check.names=FALSE)

#Skeleton
skeleton_table <-read.table("./Skeleton/Skeleton_combined_psudo_table.tsv",header=T,sep="\t",  check.names=FALSE)
skeleton_tax <-read.table("./Skeleton/Skeleton_combined_psudo_tax.tsv",header=T,sep="\t",  check.names=FALSE)
skeleton_mapping <-read.table("./Skeleton/Skeleton_combined_mapping.csv", sep=",", check.names=FALSE)

## format import files 
mucus_table <- as.data.frame(mucus_table)
class(mucus_table) <- "numeric"
as.numeric(mucus_table)
# taxa 


# create matrix format for OTU and taxonomy table
MOTU <- otu_table(as.matrix(mucus_table), taxa_are_rows = FALSE)
Mtax1 = tax_table(as.matrix(mucus_tax))

# Set metadata
MSAMPLE <- sample_data(mucus_mapping)

# Create Working phyloseq object
neutral_phylo <- phyloseq(MOTU,Mtax1,MSAMPLE,tree)
neutral_phylo


```

## Linear regression analysis
```{r}
knitr::opts_knit$set(root.dir = 'C:/Users/Colin Howe/Desktop/PSU Doctoral Program/GCMP_Working/gcmp_analysis_2024/Neutral Model Analysis')
remove(results_table)
results_table <- read.table(file = "./neutral_results_table.txt",header=T,comment.char="",sep="\t")

compartment = results_table$Compartment
as.character(compartment)
ocean = results_table$Ocean
region = results_table$Region
Rsquared = results_table$R2
faith = results_table$Coral_Host_Faiths_PD
faithsr = results_table$Faiths.SR
n_samples = results_table$n_samples
region = factor(results_table$Region)
below = results_table$n_belowMean
above = results_table$n_aboveMean
expedition = results_table$expedition
migration = results_table$m

compartment <- factor(compartment, 
                             levels = c("Mucus", "Tissue", "Skeleton"))

## Migration seems to vary significantly by expedition and host species diversity (aka habitat heterogeneous selection) but not compartments. Regional differences had less variation in migration compared to the host species diversity which has a lot of variance in migration rate with less migration in more diverse regions. A follow up would be to see if any additional regional characteristics such as temperature, algal cover etc has a linear relationship to migration.

category_migration_test <- lm(migration ~ expedition+compartment+faith)
summary(category_migration_test)

## R2 seems to vary significantly by compartment and by expedition. Compartment shows stronger structure of coral microbiome community composition. 
## Migration across compartments seems to be even but neutrality across compartments are not the same. What does this mean about how these communities are structured? Try another approach to identify if the variance in migration and R2 is greater across compartments or regions. 
category_Rquared_test <- lm(Rsquared ~ expedition+compartment+faith)
summary(category_Rquared_test) 

category_migration_test <- lm(migration ~ compartment+region+faith+faithsr+ocean)
summary(category_migration_test)

category_Rquared_test <- lm(Rsquared ~ compartment+region+faith+faithsr+ocean)
summary(category_Rquared_test) 


## Plot Linear regression
faith_m_coe <- paste0("fatihpd_pvalue:\n",round(category_migration_test$coefficients[7],7))

lm_plot = plot(faith,migration, pch = 16, cex=1.3, col="#009197", main = "Migration and Coral Faith Pd", xlab = "Coral Species Diversity (faith Pd)", ylab ="migration (m)") + text(x=1500, y=.07, label = faith_m_coe) + abline(lm(migration ~ faith), col="#EC8C5C") 

#write.table(migration_regression_results,file="migration_regression_results.txt", sep = "\t",row.names = TRUE)
#write.table(Rsquared_regression_results,file="R2_regression_results.txt", sep = "\t",row.names = TRUE)

##  Compartment Boxplot with R2 
coral_color =c("#009197","#EC8C5C","#8E3B97")

testp <- ggplot(results_table, aes(y=Rsquared, x=compartment)) + geom_boxplot(aes(color=compartment)) + theme_classic() + geom_point(aes(color=compartment)) +labs(xlab="R2", ylab="Coral Compartments", title ="Compartment Model Fit") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("#009197","#EC8C5C","#8E3B97")) + geom_signif(comparisons = list(c("Mucus", "Tissue"), c("Mucus", "Skeleton"),c("Tissue", "Skeleton")), y_position = c(.85,1, .75))

```

## Faith Pd  Analysis
```{r}
coral_object <-subset_samples(coral, BiologicalMatter == "Coral Mucus") # all corals without out groups, environmental samples & subset expeditions (i.e, singapore).
coral_object

# pull out sample data 
phyloseq::sample_data(coral_object) %>%  
    group_by(sample_name_backup, expedition_number,BiologicalMatter) %>%
  as.data.frame() %>%
  select(sample_name_backup, expedition_number,BiologicalMatter) -> microbial_faith

#Microbial Phyloseq Pd analysis 
estimate_pd(coral_object) %>%
  as.data.frame() -> microbial_faith_pd

microbial_faith$faith_pd <- microbial_faith_pd$PD
microbial_faith$faith_SR <- microbial_faith_pd$SR

# Create data frame from sample data 
phyloseq::sample_data(coral_object) %>%  
  group_by(expedition_number, BiologicalMatter,Huang_Roy_tree_name) %>%
  as.data.frame() %>%
  select(expedition_number, BiologicalMatter,Huang_Roy_tree_name)-> test_df

# Create a new column titled eco to join expedition and biological matter
test_df$eco <-paste(test_df$expedition_number,test_df$BiologicalMatter, sep = "_")

#group and count total for each unique group maintaining NA values
test_df %>% group_by(eco, Huang_Roy_tree_name) %>%  summarise(counts=n()) %>%
  ungroup %>%
  complete(nesting(eco),
           nesting(Huang_Roy_tree_name),
           fill = list(quantity = 0)) -> test_table

# Fill NA with 0
test_table[is.na(test_table)] <-0

#Build Matrix
e <- unique(test_table$eco) 
t <- unique(test_table$Huang_Roy_tree_name) 
c <- test_table$counts

test_matrix <- matrix(c, nrow = length(e), ncol = length(t), byrow=TRUE)
rownames(test_matrix) = e
colnames(test_matrix) = t

# clean data set to match each other
clean_tree <- match.phylo.comm(phy = coral_tree, comm = test_matrix)$phy
clean_comm <- match.phylo.comm(phy = coral_tree, comm = test_matrix)$comm

coral_faith_pd <- pd(clean_comm, clean_tree, include.root=TRUE)
coral_faithpd_reorded <-coral_faith_pd[order(coral_faith_pd$PD, decreasing=TRUE),] 

write.table(microbial_faith, file ="./carib_mucus_microbial_faith_pd_table.tsv", sep = ",",col.names = TRUE)

write.tsv(coral_faithpd_reorded, file ="./carib_mucus_Host_faith_pd_table.tsv" ,sep = ",",row.names = TRUE)
```

# BAR PLOTS: Coral Compartments
```{r}

##import neutral Model results tables

tissue_table <- read.table(file ="gcmp_qiime_input/Tissue_neutral_model.tsv",header=T,comment.char="", row.names=1, sep=",")
mucus_table <- read.table(file ="gcmp_qiime_input/Mucus_neutral_model.tsv",header=T,comment.char="", row.names=1, sep=",")
skeleton_table <- read.table(file ="gcmp_qiime_input/Skeleton_neutral_model.tsv",header=T,comment.char="", row.names=1, sep=",")
combined_table <- read.table(file ="gcmp_qiime_input/combined_neutral_table.tsv",header=T,comment.char="", row.names=1, sep=",")

sig_tissue <-filter(tissue_table, padj <=0.05)
sig_skeleton <-filter(skeleton_table, padj <=0.05)
sig_mucus <-filter(mucus_table, padj <=0.05)
sig_combined <-filter(combined_table, padj <=0.05)
filter(sig_combined, Phylum == "Proteobacteria") %>%
            mutate(Order= factor(Order),
                   Order= fct_reorder(Order,p, .desc = FALSE)) -> protobacteria



many_col <- c("#A6CEE3", "#33A02C","#E7298A","#FDBF6F", "#B2DF8A", "#1F78B4","#E31A1C","#6A3D9A", "#FF7F00","#1B9E77","#B15928", "#FB9A99", "#CAB2D6", "#377EB8","#D95F02", "#7570B3","#FFFF99","#666666" ,"#66A61E", "#E6AB02","#A6761D","black", "#CAB2D6",  "#B15928","#FFFF99", "#D95F02", "#E7298A", "#7570B3","#666666" , "#E6AB02", "brown3","#A6761D", "#377EB8", "#4DAF4A" ,"#984EA3", "#FFFF33","#FF7F00","cyan3", "#4DAF4A" ,"#FF7F00","#984EA3", "#A65628","#1B9E77", "#F781BF", "#F46D43","#ABDDA4", "#FDAE61","black", "#FEE08B","#E6F598","#3288BD","deeppink","#66C2A5","cyan1","red3", "#E69F00","#56B4E9","#999999","#F0E442","#0072B2","#D55E00","#D53E4F","yellow","skyblue","coral4","#A6CEE3","white","darkkhaki","brown1","#FFFF33","chocolate","darkorchid2","#FF7F00","#1B9E77","#66A61E","#FDBF6F", "#FB9A99","#A6CEE3", "#33A02C","#E7298A","#FDBF6F", "#B2DF8A", "#1F78B4","#E31A1C","#6A3D9A", "#FF7F00","#1B9E77","#B15928", "#FB9A99", "#CAB2D6", "#377EB8","#D95F02", "#7570B3","#FFFF99","#666666" ,"#66A61E", "#E6AB02")


proto_bar <- ggplot(data = protobacteria, aes(x = model, y = Order,size=p,color = Order)) +
         geom_point() + scale_color_manual(values = c(many_col))+ theme_minimal()+labs( x=" family", y="average relative abundance (p)") + guides(color="none")+ ggtitle("Non-Neutral Proteobacteria Bargraph") + theme(plot.title = element_text(hjust = 0.5))



 
  facet_wrap(~sample_type, scales= "free_x", nrow=1) +
  labs( x=" family", y="average relative abundance (p)") + guides(fill = guide_legend(title = "Order"))+ ggtitle("Non-Neutral Proteobacteria Bargraph") + scale_fill_manual(name=NULL, values = c(many_col)) 


all_bar <- ggplot(data = sig_combined, aes(x = model, y = p, fill = Phylum)) +
         geom_bar(stat = "identity", position="stack") + 
  facet_wrap(~sample_type, scales= "free_x", nrow=1) +
  labs( x=" family", y="average relative abundance (p)") + ggtitle("Compartment Neutral Bargraph") + scale_fill_manual(name=NULL, values = c(many_col)) 

mucus_bar <- ggplot(data = sig_mucus, aes(x = model, y = p, fill = Phylum)) +
         geom_bar(stat = "identity", position="stack") + 
  labs( x=" family", y="average relative abundance (p)") + ggtitle("Mucus Neutral Bargraph") + scale_fill_manual(name=NULL, values = c(many_col)) 

tissue_bar <- ggplot(data = sig_tissue, aes(x = model, y = p, fill = Phylum)) +
         geom_bar(stat = "identity", position="stack")  +
  labs( x=" family", y="average relative abundance (p)") + ggtitle("Tissue Neutral Bargraph") +  scale_fill_manual(name=NULL, values = c(many_col)) 

skeleton_bar <- ggplot(data = sig_skeleton, aes(x = model, y = p, fill = Phylum)) +
         geom_bar(stat = "identity", position="stack")  +
  labs( x=" family", y="average relative abundance (p)") + ggtitle("Skeleton Neutral Bargraph") +  scale_fill_manual(name=NULL, values = c(many_col)) 

```

# Calculate DISTANCE MATRICES
```{r}
## Mucus

bray_muc = phyloseq::distance(clean_mucus, method="bray")
uni_muc = phyloseq::distance(clean_mucus, method="unifrac")
wuni_muc= phyloseq::distance(clean_mucus, method="wUnifrac")

# Create data frame for downstream analysis
muc_df <- data.frame(sample_data(clean_mucus))


```

# Ordination from Distance Matrix and PCoA Plots 
```{r}
##Calculate the PCoA on Complex Corals
rt.pcoa = ordinate(mucus_phylo, method="PCoA", distance=uni_muc)
# Set variables to zero for subsequent changes
#Ofra
pcoa<-0
pcoa <- plot_ordination(mucus_phylo, rt.pcoa ,color="model")+  geom_point(size=3) + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + stat_ellipse(level = 0.95,type = "norm",aes(group=model))+ theme(legend.text=element_text(size=12)) + theme(axis.text=element_text(size=12, face = "bold"), axis.title=element_text(size=14,face="bold") + theme(text = element_text(size = 20,face = "bold"))) # title =expression(paste(italic(" Orbicella spp")," Bray-Curtis Distance PCoA across Depth")) scale_color_manual(values=c("#E31A1C","#1F78B4"))+
                      

```

# Alpha Diversity BoxPlots Compartments & Family
```{r, boxplots}
###Once we have our sub_samples rarefied we can start with some diversity analysis  and visualization
### Alpha Diversity ** you can make multiple diversity box plots by switch x for sample variables

complex_diversity_compartment <- plot_richness(complex, x="family", color = "Genus", title = "Merulinidae Diversity Across compartment", measures = c("Shannon","Observed"),)+
  geom_boxplot() +
  theme_classic() +
  theme(strip.background = element_blank(), plot.title = element_text(hjust = 0.5), axis.text.x.bottom = element_text(angle = -90))
 
```

## 1. PERMANOVAs Analysis
```{r}

#env material include mucus, tissue and skeleton, and outgroups without eDNA
distance_methods <-c("bray_muc","wuni_muc","uni_muc")
set.seed(129)

# This is what is called a for loop.        
for (i in distance_methods){ 
  form <- as.formula(paste(i,"model", sep="~"))
  print(form)
 adonis2(form, data=muc_df)-> result
 print(result)
}
 

## Post-hoc comparison between Coral Compartments and Location
bray_coral_pairwise <-pairwise.perm.manova(bray_coral, coral_df$country_region,nperm=1000) 
uni_coral_pairwise <-pairwise.perm.manova(uni_coral,coral_df$biological_matter,nperm=1000)
wuni_coral_pairwise <-pairwise.perm.manova(wuni_coral,coral_df$biological_matter,nperm=1000)
uni_all_com_post <-pairwise.perm.manova(uni_all,all_out_df$political_area,nperm=1000)
uni_all_bio_post <-pairwise.perm.manova(uni_all,all_out_df$biological_matter,nperm=1000)

```

# Hclutering
```{r}
## Use Distance matricies created above

# run hclust command
hc_complex_muc <- hclust(wuni_flavo, method = "ward.D2")

# Get Tree tip label names 
coral_family <- flavo_dd$BiologicalMatter

# Add tip labels to hclust object
hc_complex_muc$labels <- coral_family

## Create the Dendrogram
dend <- as.dendrogram(hc_complex_muc)

#dend <- color_branches(dend, k=9)


## clean up plot
dend %>% set("leaves_pch", 15) %>% 
  set("leaves_cex", 0.5) %>% # adjust the leave shape size
  set("labels_cex",1.0) %>% # adjust the leave label size
  color_branches(dend, k=4) %>%
  set("labels_col", k=4) %>%
  hang.dendrogram(hang_height = -0.002) %>% # hang level
  plot(main = "Caribbean Actinomarinaceae Dendrogram",
  horiz =  FALSE,  nodePar = list(cex = 0.01))


```
## Parking lot
```{r}
## Subset glom OTU table
phyloseq::otu_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_table

## Subset glom tax table
phyloseq::tax_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_tax

## Subset glom tax table
phyloseq::sample_data(glom)%>%
        as.data.frame() -> glom_metadata

# Connnect taxonomy information to neutral  model results
inner_join(neutral_results_padj,glom_tax, by = "id") %>% as.data.frame() -> nonneutral_tax


#subset otu table based on neutral results
neutral_table <-subset(glom_otu_table, id  %in%  neutral_fam_df$id)

#subset taxonomy table based on neutral results
neutral_tax <- subset(glom_otu_tax, id  %in%  neutral_fam_df$id)# & Family %in% neutral_fam_df$Family)

#subset Metadata based on neutral results
neutral_sample_name <- colnames(neutral_table)
neutral_meta <-subset(glom_metadata,sample_name_backup %in% neutral_sample_name)
neutral_metadata <- sample_data(neutral_meta)

## This approch helps transpose the data without adding numbers as colummn id
test_table <- setNames(data.frame(t(glom_otu_table[,-1])),glom_otu_table[,1])

# create matrix format for OTU and taxonomy table
OTU2 <- otu_table(data.matrix(test_table, rownames.force = NA), taxa_are_rows = TRUE)
tax2 = tax_table(as.matrix(neutral_tax))

# Create Working phyloseq object
neutral_phylo <- phyloseq(OTU2,tax2,neutral_metadata)

neutral_phylo

############################# Glom Psudo Tables ##############################################################

########### Redo for neutral without filtering for  significant padj ########
phyloseq::tax_table(glom) %>%
  as.data.frame() %>% 
  rownames_to_column("id") -> Tissue_glom_tax

## Inner_join will connect  ASV id with taxonomy using id column
inner_join(neutral_results_padj,Tissue_glom_tax, by = "id") -> non_neutral_tax

## Subset ASV from each model and signifcance 
sig_neutral_tax = select(filter(non_neutral_tax, model == "neutral"),id,p_abundance,freq,padj,Phylum,Class,Order,Family,model)

## Subset glom OTU table
phyloseq::otu_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_table

##subset otu table based on neutral results and rename column names
neutral_table <-subset(glom_otu_table, id  %in%  sig_neutral_tax$id)
n <-length(neutral_table)
colnames(neutral_table)[2:n] <- paste(colnames(neutral_table)[2:n], "neutral", sep = "_")

## Subset metadata file
data.frame(sample_data(glom)) %>% select(sample_name_backup, expedition_number,host_scientific_name) %>% as.data.frame() -> glom_meta

## Create new columns for the metadata data
glom_meta %>% mutate(psuedo_sample_name = paste(colnames(neutral_table)[2:n]),
                         compartment_neutraility = paste("Tissue", "neutral", sep = '_'),
                         compartment = paste ("Tissue"),
                         neutraility = paste("neutral")) %>% 
                         as.data.frame() -> glom_metadata

psuedo_table_name <- paste0("Tissue","_","neutral_table.csv")
(write.csv(neutral_table, file =psuedo_table_name, row.names = FALSE))

psuedo_map_name <- paste0("Tissue","_","neutral_metadata.csv")
(write.csv(glom_metadata, file =psuedo_map_name, row.names = FALSE))

psuedo_tax_name <- paste0("Tissue","_","neutral_taxonomy.csv")
(write.csv(subset_tax, file =psuedo_tax_name, row.names = FALSE))
# Subset Taxonomy and join with neutral model results
neutral = c("above","below")
for (i in neutral){
phyloseq::tax_table(glom) %>%
  as.data.frame() %>% 
  rownames_to_column("id") -> glom_tax
## Inner_join will connect  ASV id with taxonomy using id column
inner_join(neutral_results_padj,glom_tax, by = "id") -> non_neutral_tax

## Subset ASV from each model and signifcance 
sig_neutral_tax = select(filter(non_neutral_tax, model == i & padj <=0.05),id,p_abundance,freq,padj,Phylum,Class,Order,Family,model)

## Subset glom OTU table
phyloseq::otu_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_table

#subset otu table based on neutral results and rename column names
neutral_table <-subset(glom_otu_table, id  %in%  sig_neutral_tax$id)
l = length(neutral_table)
colnames(neutral_table)[2:l] <- paste(colnames(neutral_table)[2:l], i, sep = "_")

## Subset metadata file
data.frame(sample_data(glom)) %>% select(sample_name_backup, expedition_number,host_scientific_name) %>% as.data.frame() -> glom_meta
## Create new columns for the metadata data
glom_meta %>% mutate(psuedo_sample_name = paste(colnames(neutral_table)[2:299]),
                         compartment_neutraility = paste("mucus", i, sep = '_'),
                         compartment = paste ("mucus"),
                         neutraility = paste(i)) %>% 
                         as.data.frame() -> glom_metadata

## Subset significant results with taxonomy to retain only significant taxonomy
subset_tax <- subset(glom_tax, id %in% sig_neutral_tax$id) %>% 
  select(-c("Genus","Species"))

psuedo_table_name <- paste0("mucus","_",i,"_table.tsv")
(write.table(neutral_table, file =psuedo_table_name, row.names = FALSE))

psuedo_map_name <- paste0("mucus","_",i,"_metadata.tsv")
(write.table(glom_metadata, file =psuedo_map_name, row.names = FALSE))

psuedo_tax_name <- paste0("mucus","_",i,"_taxonomy.tsv")
(write.table(subset_tax, file =psuedo_tax_name, row.names = FALSE))

}

########### Redo for neutral without filtering for  significant padj ########
phyloseq::tax_table(glom) %>%
  as.data.frame() %>% 
  rownames_to_column("id") -> glom_tax

## Inner_join will connect  ASV id with taxonomy using id column
inner_join(neutral_results_padj,glom_tax, by = "id") -> non_neutral_tax

## Subset ASV from each model and signifcance 
sig_neutral_tax = select(filter(non_neutral_tax, model == "neutral"),id,p_abundance,freq,padj,Phylum,Class,Order,Family,model)

## Subset glom OTU table
phyloseq::otu_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_table

##subset otu table based on neutral results and rename column names
neutral_table <-subset(glom_otu_table, id  %in%  sig_neutral_tax$id)
colnames(neutral_table)[2:299] <- paste(colnames(neutral_table)[2:299], "neutral", sep = "_")

## Subset metadata file
data.frame(sample_data(glom)) %>% select(sample_name_backup, expedition_number,host_scientific_name) %>% as.data.frame() -> glom_meta

## Create new columns for the metadata data
glom_meta %>% mutate(psuedo_sample_name = paste(colnames(neutral_table)[2:299]),
                         compartment_neutraility = paste("mucus", "neutral", sep = '_'),
                         compartment = paste ("mucus"),
                         neutraility = paste("neutral")) %>% 
                         as.data.frame() -> glom_metadata

psuedo_table_name <- paste0("mucus","_","neutral_table.tsv")
(write.table(neutral_table, file =psuedo_table_name, row.names = FALSE))

psuedo_map_name <- paste0("mucus","_","neutral_metadata.tsv")
(write.table(glom_metadata, file =psuedo_map_name, row.names = FALSE))

psuedo_tax_name <- paste0("mucus","_","neutral_taxonomy.tsv")
(write.table(subset_tax, file =psuedo_tax_name, row.names = FALSE))




```

```