---
title: "Global Coral Microbiome Project: Statistical Analysis and Modelling Workflow 2024"
authors: "Colin Howe, Sofia Roitman"
start date: "01/17/2024"
current date: "04/21/2024"
output: html_document
---

# First, Set working Directory load library and Import Data into Phyloseq
```{r warning=FALSE}
# Command for setting your work directory
setwd("C:/Users/Colin Howe/Desktop/PSU Doctoral Program/GCMP_Working/gcmp_analysis_2024")

# Print your working directory
#getwd() 

```

## Load required libraries. 
```{r, message=FALSE, warning=FALSE} 
library(BiocManager)
library(phyloseq)
library(dplyr)
library(DESeq2)
library(ComplexHeatmap)
library(ade4)
library(vegan)
library(RColorBrewer)
library(ggplot2)
library(ape)
library(microbiome)
library(dendextend)
library(gplots)
library(ggrepel)
library(tidyr)
library(ggpubr)
library(cowplot)
library(RVAideMemoire)
library(microbiomeMarker)
library(qiime2R)
library(tidyverse)

```

# Import OTU Table Taxonomy and Tree from Qiime2 and import Metadata
```{r}

#Import from .qza file into a phyloseq object
asv <- qza_to_phyloseq(features = "carib_silva_merged_table.qza")

#### Import Metadata read.table
metadata <- read.table(file = "carib_merged_mapping.csv",header=T,comment.char="", row.names=1, sep=",")

### Import Tree file from biom output tree.nwk
tree2 <- read_tree("tree.nwk")

### Import taxonomy from biom output as .tsv format using read.table
taxonomy <- read.table(file = "taxonomy.tsv", sep = "\t", header = T ,row.names = 1)

# clean the taxonomy

##code referenced from Yan Hui: email me@yanh.org github: yanhui09
tax <- taxonomy %>%
  select(Taxon) %>% 
  separate(Taxon, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), "; ")

tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "k__",""),
                        Phylum = str_replace(tax[,2], "p__",""),
                        Class = str_replace(tax[,3], "c__",""),
                        Order = str_replace(tax[,4], "o__",""),
                        Family = str_replace(tax[,5], "f__",""),
                        Genus = str_replace(tax[,6], "g__",""),
                        Species = str_replace(tax[,7], "s__",""),
                        stringsAsFactors = FALSE)

tax.clean[is.na(tax.clean)] <- ""
tax.clean[tax.clean=="__"] <- ""

for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,7] != ""){
    tax.clean$Species[i] <- paste(tax.clean$Genus[i], tax.clean$Species[i], sep = " ")
  } else if (tax.clean[i,2] == ""){
    kingdom <- paste("Unclassified", tax.clean[i,1], sep = " ")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Unclassified", tax.clean[i,2], sep = " ")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Unclassified", tax.clean[i,3], sep = " ")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Unclassified", tax.clean[i,4], sep = " ")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Unclassified", tax.clean[i,5], sep = " ")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Unclassified ",tax.clean$Genus[i], sep = " ")
  }
}

# create matrix format for OTU and taxonomy table
OTU <- otu_table(as.matrix(asv), taxa_are_rows = TRUE)
tax1 = tax_table(as.matrix(tax.clean))

# Set metadata
SAMPLE <- sample_data(metadata)

# Read tree.nwk
TREE = read_tree(tree2)

# Create Working phyloseq object
main_coral <- phyloseq(OTU,tax1,SAMPLE,TREE)

# visualize tax table
table(tax_table(main_coral)[,"Kingdom"])

# remove unknown bacteria or unassigned
phylo_noking <-main_coral %>%
  phyloseq::subset_taxa(!Kingdom %in% c("Unassigned","Unclassified d__Bacteria","d__Eukaryota")) %>%
  phyloseq::subset_taxa(!Phylum %in% c("Unclassified Unassigned","Unclassified d__Archaea","Unclassified d__Bacteria","Unclassified d__Eukaryota")) %>%
  phyloseq::subset_taxa(!Family %in% c("Mitochondria","Chloroplast"))

# Remove any samples with less than 1 read
phylo_coral = prune_samples(sample_sums(phylo_noking)>1, phylo_noking)
phylo_coral

# Subset only corals from database
coral <-subset_samples(phylo_coral, sample_type_EMP == "coral" & outgroup == "n" & within_group == "yes") # all corals without out groups, environmental samples & subset expeditions (i.e, singapore).
coral
## Check: 
table(tax_table(coral)[,"Family"])

coral.rarefied = rarefy_even_depth(coral, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
coral.rarefied
##Agglomerate taxa to family 
glom_coral <- tax_glom(coral.rarefied, taxrank = 'Family', NArm = TRUE) ## 444 Samples 
coral
coral.rarefied
glom_coral

phyloseq::otu_table(coral.rarefied)%>%
        as.data.frame() %>%
        rownames_to_column("id") %>%
        left_join(phyloseq::tax_table(coral.rarefied) %>%
        as.data.frame() %>%
        rownames_to_column("id")) -> coral_biom
        

write.csv(coral_biom, file ="GCMP24_rarefied_coral.csv")
print(paste("Generating Rarefied Coral dataset:", coral.rarefied))

```

### Subset Caribbean Phyloseq Objects for Comparative Analysis 
```{r}
## Caribbean dataset rarefication to 2000 per sample for downstream analysis Retained 1,222,000 (22.55%) features in 611 (82.12%) samples at the specifed sampling depth.

#colombia5 <-subset_samples(coral, political_area == "Colombia" & expedition_number == "E5" & sample_type_EMP == "coral" & outgroup == "n")
#colombia5

# rarefy to even depth
#colombia.rarefied5 = rarefy_even_depth(colombia5, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#colombia.rarefied5

#### Subsample Colombia Expedition E12 This expedition is mainly comprised of sequences from 2022
colombia12 <-subset_samples(coral, political_area == "Colombia" & expedition_number == "E12" & sample_type_EMP == "coral"& outgroup == "n")
colombia12

# rarefy to even depth
colombia.rarefied12 = rarefy_even_depth(colombia12, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
colombia.rarefied12

#### Subsample Corals from Colombia includes E5 & E12
colombia <-subset_samples(coral, political_area == "Colombia" & sample_type_EMP == "coral" & outgroup == "n")
colombia

# rarefy to even depth
colombia.rarefied = rarefy_even_depth(colombia, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
colombia.rarefied

#### Subsample Orbicella from Colombia E5 & E12
#col_orb <-subset_samples(coral, political_area == "Colombia" & host_genus_id == "Orbicella")# political_area == "Colombia" & expedition_number == "E5")
#col_orb

# rarefy to even depth
#col.orb.rarefied = rarefy_even_depth(col_orb, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#col.orb.rarefied

#### Subsample Orbicella faveolata E5 & E12
#faveolata <-subset_samples(coral, political_area == "Colombia" & host_species == "faveolata")
#faveolata

# rarefy to even depth faveolata
#faveolata.rarefied = rarefy_even_depth(faveolata, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#faveolata.rarefied


############# Subsample Panama political_area including E2 & E11 East & West################
#panama <-subset_samples(coral, political_area == "Panama" & sample_type_EMP == "coral" & outgroup == "n")
#panama

# rarefy to even depth
#panama.rarefied = rarefy_even_depth(panama, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#panama.rarefied

#### Subsample Panama west E11
panama.w <-subset_samples(coral, political_area == "Panama" & expedition_number == "E11" & sample_type_EMP == "coral" & outgroup == "n")
panama.w

# rarefy to even depth
panama.w.rarefied = rarefy_even_depth(panama.w, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
panama.w.rarefied

#### Subsample Panama east E2
panama.e <-subset_samples(coral, political_area == "Panama" & expedition_number == "E2" & sample_type_EMP == "coral" & outgroup == "n")
panama.e

# rarefy to even depth
panama.e.rarefied = rarefy_even_depth(panama.e, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
panama.e.rarefied

#### Subsample Panama Orbicella Corals
#pan_orb <-subset_samples(coral, political_area == "Panama" & host_genus_id == "Orbicella")
#pan_orb

# rarefy to even depth
#pan.orb.rarefied = rarefy_even_depth(pan_orb, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#pan.orb.rarefied

#### Subsample Panama Orbicella faveolata Corals
#pan_ofav <-subset_samples(coral, political_area == "Panama" & host_species == "faveolata")
#pan_ofav

# rarefy to even depth
#pan.ofav.rarefied = rarefy_even_depth(pan_ofav, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#pan.ofav.rarefied


############# Subsample Orbicella and Environmental Samples################
# subset Orbicella genus across colombia and panama (all orbicella samples) expedition 5 is well mixed between 2022 and 2023 sequencing. E12 is mainly from 2022 and all orbicella corals from E12 are from 2022. 
#orbicella <- subset_samples(coral, host_genus_id == "Orbicella" & expedition_number == "E12") 

# rarefy to even depth
#orb.rarefied = rarefy_even_depth(orbicella, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#orb.rarefied

## Overall Caribbean
#carib <-subset_samples(coral, ocean_area == "Caribbean" & sample_type_EMP == "coral" & outgroup == "n")
#carib

# rarefy to even depth
#carib.rarefied = rarefy_even_depth(carib, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#carib.rarefied

#### Environmental Samples 
#env_sample <- subset_samples(coral, env_sample == "yes" & political_area == "Colombia")
#env_sample

# keep result reproductive
#env.rarefied = rarefy_even_depth(env_sample, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
#env.rarefied

#### Water 
water <- subset_samples(phylo_coral, sample_type == "Reef Water")# & political_area == "Colombia")
water

# keep result reproductive
water.rarefied = rarefy_even_depth(water, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
water.rarefied

#### Sediment 
sediment <- subset_samples(phylo_coral, sample_type == "Reef Sediment" & political_area == "Colombia")
sediment

# keep result reproductive
sed.rarefied = rarefy_even_depth(sediment, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
sed.rarefied

## Hawaii 
#### Subsample Panama east E2
hawaii <-subset_samples(coral, expedition_number == "E15" & sample_type_EMP == "coral" & outgroup == "n")
hawaii

# rarefy to even depth
hawaii.rarefied = rarefy_even_depth(hawaii, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
hawaii.rarefied

## Compartment
#### Subsample Mucus from Colombia
mucus <-subset_samples(coral,BiologicalMatter == "Coral Mucus")
mucus

# rarefy to even depth
mucus.rarefied = rarefy_even_depth(mucus, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
mucus.rarefied

#### Subsample Panama east E12 E5
tissue <-subset_samples(coral,BiologicalMatter == "Coral Tissue")
tissue
# rarefy to even depth
tissue.rarefied = rarefy_even_depth(tissue, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
tissue.rarefied

#### Subsample Panama east E12 E5
skeleton <-subset_samples(coral, BiologicalMatter == "Coral Skeleton")
skeleton
# rarefy to even depth
skeleton.rarefied = rarefy_even_depth(skeleton, rngseed=111, sample.size=2000, replace=F, trimOTUs = TRUE)
skeleton.rarefied

## Aglomerate Taxa to Family Level
glom_coral <- tax_glom(coral.rarefied, taxrank = 'Family', NArm = TRUE) ## 444 Samples 
#glom_col <- tax_glom(colombia.rarefied, taxrank = 'Family', NArm = TRUE)##  150 samples
#glom_col5 <- tax_glom(colombia.rarefied5, taxrank = 'Family', NArm = TRUE)## 65 samples
glom_col12 <- tax_glom(colombia.rarefied12, taxrank = 'Family', NArm = TRUE)## 85 samples
#glom_col_orb <- tax_glom(col.orb.rarefied, taxrank = 'Family', NArm = TRUE)## 33 samples
#glom_col12_orb <- tax_glom(orb.rarefied, taxrank = 'Family', NArm = TRUE)## 18 Samples
#glom_ofav <- tax_glom(faveolata.rarefied, taxrank = 'Family', NArm = TRUE)## 15 Samples all from Colombia
#glom_pan_ofav <- tax_glom(pan.ofav.rarefied, taxrank = 'Family', NArm = TRUE) # 9 Samples
#glom_carib <-tax_glom(carib.rarefied, taxrank = 'Family', NArm = TRUE) ## 253 from  E2 Panama and Colombia E5 & E12

#glom_pan <- tax_glom(panama.rarefied, taxrank = 'Family', NArm = TRUE)##  185 samples from Panama E2 and E11 
glom_panw <- tax_glom(panama.w.rarefied, taxrank = 'Family', NArm = TRUE)## 82 samples
glom_pane <- tax_glom(panama.e.rarefied, taxrank = 'Family', NArm = TRUE)##  102 samples
#glom_pan_orb <- tax_glom(pan.orb.rarefied, taxrank = 'Family', NArm = TRUE)## 35 samples
#glom_pan_ofav <- tax_glom(pan.ofav.rarefied, taxrank = 'Family',  NArm = TRUE)## 68 samples from the Caribbean
#glom_env <- tax_glom(env.rarefied, taxrank = 'Family', NArm = TRUE)## 52 samples from Colombia, Panama only has one environmental sample
glom_wat <- tax_glom(water.rarefied, taxrank = 'Family', NArm = TRUE)## 34 samples from Colombia, Panama only has one environmental sample
glom_sed <- tax_glom(sed.rarefied, taxrank = 'Family', NArm = TRUE)## 17 samples from Colombia, Panama only has one environmental 
glom_haw <- tax_glom(hawaii.rarefied, taxrank = 'Family', NArm = TRUE) ## 51 Samples 
glom_mucus <- tax_glom(mucus.rarefied, taxrank = 'Family', NArm = TRUE)
glom_tissue <- tax_glom(tissue.rarefied, taxrank = 'Family', NArm = TRUE)
glom_skeleton <- tax_glom(skeleton.rarefied, taxrank = 'Family', NArm = TRUE)


```

#Bruns et al. 2016 Neutral Model functions
```{r}
################define sncm.fit############
sncm.fit <- function(spp, pool=NULL, stats=TRUE, taxon=NULL){
	require(minpack.lm)
	require(Hmisc)
	require(stats4)
	
	options(warn=-1)

	#Calculate the number of individuals per community
	N <- mean(apply(spp, 1, sum))
	
	#Calculate the average relative abundance of each taxa across communities
	if(is.null(pool)){
		p.m <- apply(spp, 2, mean)
		p.m <- p.m[p.m != 0]
		p <- p.m/N
	} else {
		p.m <- apply(pool, 2, mean)
		p.m <- p.m[p.m != 0]
		p <- p.m/N
	}

#Calculate the occurrence frequency of each taxa across communities
spp.bi <- 1*(spp>0)
freq <- apply(spp.bi, 2, mean)
freq <- freq[freq != 0]
#Combine
C <- merge(p, freq, by=0)
C <- C[order(C[,2]),]
C <- as.data.frame(C)
C.0 <- C[!(apply(C, 1, function(y) any(y == 0))),] #Removes rows with any zero (absent in either source pool or local communities)
p <- C.0[,2]
freq <- C.0[,3]
names(p) <- C.0[,1]
names(freq) <- C.0[,1]

#Calculate the limit of detection
d = 1/N

##Fit model parameter m (or Nm) using Non-linear least squares (NLS)
m.fit <- nlsLM(freq ~ pbeta(d, N*m*p, N*m*(1-p), lower.tail=FALSE), start=list(m=0.1))
m.ci <- confint(m.fit, 'm', level=0.95)
	
##Fit neutral model parameter m (or Nm) using Maximum likelihood estimation (MLE)
sncm.LL <- function(m, sigma){
	R = freq - pbeta(d, N*m*p, N*m*(1-p), lower.tail=FALSE)
	R = dnorm(R, 0, sigma)
	-sum(log(R))
}
	
m.mle <- mle(sncm.LL, start=list(m=0.1, sigma=0.1), nobs=length(p),"Nelder-Mead")

##Calculate Akaike's Information Criterion (AIC)
aic.fit <- AIC(m.mle, k=2)
bic.fit <- BIC(m.mle)

##Calculate goodness-of-fit (R-squared and Root Mean Squared Error)
freq.pred <- pbeta(d, N*coef(m.fit)*p, N*coef(m.fit)*(1-p), lower.tail=FALSE)
Rsqr <- 1 - (sum((freq - freq.pred)^2))/(sum((freq - mean(freq))^2))
RMSE <- sqrt(sum((freq-freq.pred)^2)/(length(freq)-1))
	
pred.ci <- binconf(freq.pred*nrow(spp), nrow(spp), alpha=0.01, method="wilson", return.df=TRUE)
	
##Calculate AIC for binomial model
bino.LL <- function(mu, sigma){
R = freq - pbinom(d, N, p, lower.tail=FALSE)
R = dnorm(R, mu, sigma)
-sum(log(R))
	}
bino.mle <- mle(bino.LL, start=list(mu=0, sigma=0.1), nobs=length(p))
	
aic.bino <- AIC(bino.mle, k=2)
bic.bino <- BIC(bino.mle)
	
##Goodness of fit for binomial model
bino.pred <- pbinom(d, N, p, lower.tail=FALSE)
Rsqr.bino <- 1 - (sum((freq - bino.pred)^2))/(sum((freq - mean(freq))^2))
RMSE.bino <- sqrt(sum((freq - bino.pred)^2)/(length(freq) - 1))

bino.pred.ci <- binconf(bino.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)
	
##Calculate AIC for Poisson model
pois.LL <- function(mu, sigma){
	R = freq - ppois(d, N*p, lower.tail=FALSE)
	R = dnorm(R, mu, sigma)
	-sum(log(R))
}
pois.mle <- mle(pois.LL, start=list(mu=0, sigma=0.1), nobs=length(p))
	
aic.pois <- AIC(pois.mle, k=2)
bic.pois <- BIC(pois.mle)
	
##Goodness of fit for Poisson model
pois.pred <- ppois(d, N*p, lower.tail=FALSE)
Rsqr.pois <- 1 - (sum((freq - pois.pred)^2))/(sum((freq - mean(freq))^2))
RMSE.pois <- sqrt(sum((freq - pois.pred)^2)/(length(freq) - 1))

pois.pred.ci <- binconf(pois.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)

##Results
	if(stats==TRUE){
		fitstats <- data.frame(m=numeric(), m.ci=numeric(), m.mle=numeric(), maxLL=numeric(), binoLL=numeric(), poisLL=numeric(), Rsqr=numeric(), Rsqr.bino=numeric(), Rsqr.pois=numeric(), RMSE=numeric(), RMSE.bino=numeric(), RMSE.pois=numeric(), AIC=numeric(), BIC=numeric(), AIC.bino=numeric(), BIC.bino=numeric(), AIC.pois=numeric(), BIC.pois=numeric(), N=numeric(), Samples=numeric(), Richness=numeric(), Detect=numeric())
		fitstats[1,] <- c(coef(m.fit), coef(m.fit)-m.ci[1], m.mle@coef['m'], m.mle@details$value, bino.mle@details$value, pois.mle@details$value, Rsqr, Rsqr.bino, Rsqr.pois, RMSE, RMSE.bino, RMSE.pois, aic.fit, bic.fit, aic.bino, bic.bino, aic.pois, bic.pois, N, nrow(spp), length(p), d)
		return(fitstats)
	} else {
	A <- cbind(p, freq, freq.pred, pred.ci[,2:3], bino.pred, bino.pred.ci[,2:3])
	A <- as.data.frame(A)
		colnames(A) <- c('p', 'freq', 'freq.pred', 'pred.lwr', 'pred.upr', 'bino.pred', 'bino.lwr', 'bino.upr')
		if(is.null(taxon)){
			B <- A[order(A[,1]),]
		} else {
			B <- merge(A, taxon, by=0, all=TRUE)
			row.names(B) <- B[,1]
			B <- B[,-1]
			B <- B[order(B[,1]),]
		}
		return(B)
	}
}
##Identify above, below and neutral OTUs 

addf = function(freq, pred.upr, pred.lwr) {
  if(freq > pred.upr){
    model = "above"
  }
  else if(freq < pred.lwr){
    model = "below"
  }
  else if(pred.upr >= freq & freq >= pred.lwr){
    model = "neutral"
  }
  return(model)
}

```

## Neutral Model test
```{r}
###coral communities
coral <- 0
coral <- prune_taxa(taxa_sums(glom_coral) > 0, glom_coral)
coral
otu_coral = as.data.frame(t(otu_table(coral)))

## Neutral fit
mod_w_coral = sncm.fit(otu_coral)
mod1_w_coral = sncm.fit(otu_coral,stats = F)

#min(sample_sums(glom_col_orb))
print(length(mod_w_coral$freq))

mod1_w_coral$model = mapply(addf, mod1_w_coral$freq, mod1_w_coral$pred.upr, mod1_w_coral$pred.lwr)  
length(which(mod1_w_coral$model == "above"))
length(which(mod1_w_coral$model == "below"))
length(which(mod1_w_coral$model == "neutral"))

mod1_w_coral$model <- factor(mod1_w_coral$model, 
                             levels = c("above", "neutral", "below"))

## Assign Taxonomy to table of significant microbes
coral_taxa = cbind(as(mod1_w_coral, "data.frame"), as(tax_table(glom_coral)[rownames(mod1_w_coral), ], "matrix"))
#coral_taxa
write.csv(coral_taxa,"Corals from E2 E5 E11 E12 E15 Neutral Taxonomy table.csv",row.names = TRUE)


set.seed(111)
###seawater communities
#seawater <- subset_samples(physeq2, SampleType == "seawater")
seawater <- prune_taxa(taxa_sums(glom_wat) > 0, glom_wat )
seawater

otu_seawater = as.data.frame(t(otu_table(seawater)))

mod_w_seawater = sncm.fit(otu_seawater)
mod1_w_seawater = sncm.fit(otu_seawater,stats = F)


mod1_w_seawater$model = mapply(addf, mod1_w_seawater$freq, mod1_w_seawater$pred.upr, mod1_w_seawater$pred.lwr)  
length(which(mod1_w_seawater$model == "above"))
length(which(mod1_w_seawater$model == "below"))
length(which(mod1_w_seawater$model == "neutral"))

mod1_w_seawater$model <- factor(mod1_w_seawater$model, 
                                levels = c("above", "neutral", "below"))

## Assign Taxonomy to table of significant microbes
env_taxa = cbind(as(mod1_w_seawater, "data.frame"), as(tax_table(glom_wat)[rownames(mod1_w_seawater), ], "matrix"))

write.csv(env_taxa,"Colombia Water Samples Neutral Taxonomy table.csv",row.names = TRUE)


```

## Plot Neutral Model
```{r}
#plot neuModel
p_neuM_coral=ggplot(mod1_w_coral, aes(x=log10(p), y=freq, color = model))+
  geom_point(shape=19, alpha=0.7, size=2.5)+
  geom_line(aes(x=log10(p), y=freq.pred),
            color="blue",linetype="solid",linewidth=1)+
  geom_line(aes(x=log10(p), y=pred.lwr),
            color="blue",linetype="dashed",linewidth=1)+
  geom_line(aes(x=log10(p), y=pred.upr),
            color="blue",linetype="dashed",linewidth=1)+
    scale_colour_manual(name="OTUs",
                      values=c("#E41A1C", "#999999", "#FF7F00"),
                      breaks=c("above", "neutral", "below"),
                      labels=c("Over-represented", "Neutrally-distributed", "Under-represented"))+
  xlab("Log10 (Mean Relative abundance)") + ylab("Occurance frequency")+
  theme_bw() +
  theme(axis.title.x = element_text(face="bold", size=15),
        axis.title.y = element_text(face="bold", size=15),
        axis.text.x = element_text(colour = "black", size=12),
        axis.text.y = element_text(colour = "black", size=12), 
        legend.title = element_text(size=11.5, face = "bold"),
        legend.text = element_text(size=11.5),
        legend.position="none")


lb1 <- mod_w_coral$Rsqr
lb2 <- mod_w_coral$m
lb3 <- expression(paste("Corals from E2 E5 E11 E12 E15 expeditions "))

p_neuM_coral1 = p_neuM_coral + annotate("text", x = -3.5, y=0.7, size=6.5, label=lb1, 
                                        fontface="bold", color="black",
                                        parse=TRUE) +
  annotate("text", x = -3.5, y=0.6, size=6.5, label=lb2, 
           fontface="bold", color="black",
           parse=TRUE) +
  annotate("text", x = -3.5, y=0.9, size=6.5, label=lb3, 
           fontface="bold", color="black",
           parse=TRUE)


#plot neuModel
p_neuM_seawater=ggplot(mod1_w_seawater, aes(x=log10(p), y=freq, color = model))+
  geom_point(shape=19, alpha=0.7, size=2.5)+
  geom_line(aes(x=log10(p), y=freq.pred),
            color="blue",linetype="solid",size=1)+
  geom_line(aes(x=log10(p), y=pred.lwr),
            color="blue",linetype="dashed",size=1)+
  geom_line(aes(x=log10(p), y=pred.upr),
            color="blue",linetype="dashed",size=1)+
  scale_colour_manual(name="OTUs",
                      values=c("#E41A1C", "#999999", "#FF7F00"),
                      breaks=c("above", "neutral", "below"),
                      labels=c("Over-represented", "Neutrally-distributed", "Under-represented"))+
  xlab("Log10 (Mean Relative abundance)") + ylab("Occurance frequency")+
  theme_bw() +
  theme(axis.title.x = element_text(face="bold", size=15),
        axis.title.y = element_text(face="bold", size=15),
        axis.text.x = element_text(colour = "black", size=12),
        axis.text.y = element_text(colour = "black", size=12), 
        legend.title = element_text(size=11.5, face = "bold"),
        legend.text = element_text(size=11.5),
        legend.position="none")#http://127.0.0.1:24155/graphics/plot_zoom_png?width=1762&height=934

ev1 <- mod_w_seawater$Rsqr
ev2 <- mod_w_seawater$m
ev3 <- expression(paste("Colombian Water  Samples"))

p_neuM_seawater1 = p_neuM_seawater + annotate("text", x = -3.5, y=0.7, size=6.5, label=ev1, 
                                              fontface="bold", color="black",
                                              parse=TRUE) +
annotate("text", x = -3.5, y=0.6, size=6.5, label=ev2, 
           fontface="bold", color="black",
           parse=TRUE) +
annotate("text", x = -3.5, y=0.9, size=5.5, label=ev3, 
           fontface="bold", color="black",
           parse=TRUE)

getwd()

mp_ncm = ggdraw() +
  draw_plot(p_neuM_coral1, x = 0.01, y = 0.1, width = 0.49, height = 0.8) + 
  draw_plot(p_neuM_seawater1, x = 0.51, y = 0.1, width = 0.49, height = 0.8) 
mp_ncm
#ggsave("neuModel.jpg", mp_ncm, width = 20, height = 10, dpi= 600)


```

# Calculate DISTANCE MATRICES
```{r}
## Distance Matrices
# remove outlier
to_remove <- ("230-HI62-T")
orb_prune <- prune_samples(!(sample_names(coral.rarefied) %in% to_remove), coral.rarefied)
## All Corals 
bray_coral = phyloseq::distance(orb_prune, method="bray")
uni_coral = phyloseq::distance(orb_prune, method="unifrac")
wuni_coral = phyloseq::distance(orb_prune, method="wUnifrac")

# Create data frame for downstream analysis
coral_df <- data.frame(sample_data(orb_prune))

```

# Ordination from Distance Matrix and PCoA Plots 
```{r}

##Calculate the PCoA on Complex Corals
rt.pcoa = ordinate(coral.rarefied, method="PCoA", distance=uni_coral)
# Set variables to zero for subsequent changes
# expedition_number
pcoa<-0
pcoa <- plot_ordination(coral.rarefied, rt.pcoa,  shape = "sample_type",color="family")+  geom_point(size=3) + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + stat_ellipse(level = 0.95,type = "norm",aes(group=family))+ theme(legend.text=element_text(size=12)) + theme(axis.text=element_text(size=12, face = "bold"), axis.title=element_text(size=14,face="bold") + theme(text = element_text(size = 20,face = "bold")))
                      

```

## 1. PERMANOVAs Analysis
```{r}

#env material include mucus, tissue and skeleton, and outgroups without eDNA
distance_methods <-c("bray_coral","uni_coral","wuni_coral")
set.seed(129)

# This is what is called a for loop.        
for (i in distance_methods){ 
  form <- as.formula(paste(i,"expedition_number+family+sample_type+Huang_Roy_tree_name+host_genus_id", sep="~"))
  print(form)
 adonis2(form, data=coral_df)-> result
 print(result)
}
 

## Post-hoc comparison between Coral Compartments and Location

bray_carib_pairwise <-pairwise.perm.manova(bray_coral, coral_df$family,nperm=1000,p.method = "BH") 
uni_carib_pairwise <-pairwise.perm.manova(uni_coral,coral_df$family,nperm=1000,p.method = "BH")
wuni_carib_pairwise <-pairwise.perm.manova(wuni_coral,coral_df$family,nperm=1000,p.method = "BH")

```

# Alpha Diversity BoxPlots Compartments & Family
```{r, boxplots}
###Once we have our sub_samples rarefied we can start with some diversity analysis  and visualization
### Alpha Diversity ** you can make multiple diversity box plots by switch x for sample variables

complex_diversity_compartment <- plot_richness(coral, x="expedition_number", color = "sample_type", title = "Coral Diversity Across compartment", measures = c("Shannon","Observed"),)+
  geom_boxplot() +
  theme_classic() +
  theme(strip.background = element_blank(), plot.title = element_text(hjust = 0.5), axis.text.x.bottom = element_text(angle = -90))
 
```

# BAR PLOTS: Coral Compartments
```{r, message=FALSE, warning=FALSE}
## Flavobacteriaceae & Cyclobacteriaceae seems to be the most frequent and abundant between Pamana east & West
## Found Across Ofav samples in Colombia E5: Order: Halanaerobiales Family: Halanaerobiaceae
## Halobacteroidaceae
## Colombia Neutral Model shows Actinomarinaceae high relative abundance but low frequency
#Flavobacteriaceae, Cyanobiaceae, Rhodobacteraceae, Endozoicomonadaceae, Amoebophilaceae
#Nitrosopumilaceae, Cyclobacteriaceae, Sphingomonadaceae, Alteromonadaceae
## This was the most abundance across the Caribbean: Fokiniaceae

#### Use from agglomerated phyloseq objects from Neutral Model Section
flavo <- subset_taxa(coral.rarefied, Family == "Rhodobacteraceae", NArm = TRUE)
flavo_prune = prune_samples(sample_sums(flavo)>1, flavo)
# Transform sample counts to proportion out of 100

glom <- tax_glom(flavo, taxrank = 'Genus', NArm = TRUE)

ps.all = transform_sample_counts(flavo_prune, function(x) x/sum(x)*100)
ps.melt <- psmelt(ps.all)

# change to character for easy-adjusted level
ps.melt <- ps.melt %>%
  group_by(Genus,BiologicalMatter,expedition_number,site_name) %>%
  mutate(median=median(Abundance)) %>%
  arrange(desc(median))

# select group median > 1
keep <- unique(ps.melt$Genus[ps.melt$median >= 1])
ps.melt$Genus[!(ps.melt$Genus %in% keep)] <- "< %1 (Other)"

#to get the same rows together
#to get the same rows together
ps.melt_sum <- ps.melt %>%
  group_by(Genus,BiologicalMatter,expedition_number,site_name) %>%
  summarise(Abundance = sum(Abundance), 
            mean = sum(Abundance), .groups = "drop") %>%
            mutate(Genus= factor(Genus),
                   Genus= fct_reorder(Genus ,mean, .desc = TRUE),
                   Genus= fct_shift(Genus, n=1))

# create an object with color blind safe orientation
many_col <- c("#A6CEE3", "#33A02C", "#B2DF8A",  "#1F78B4","#E31A1C","#6A3D9A", "#FF7F00","#FDBF6F", "#FB9A99", "#CAB2D6",  "#B15928","#1B9E77", "#D95F02", "#FFFF99","#E7298A", "#7570B3","#666666" ,"#66A61E", "#E6AB02","#A6761D", "#E41A1C", "#377EB8", "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF" ,"#D53E4F","#F46D43", "#FDAE61", "#FEE08B","#999999", "#E6F598","#ABDDA4","#3288BD","cyan1","#66C2A5")

testp <-0
testp <-ggplot(ps.melt_sum, aes(x = BiologicalMatter, y = Abundance, fill = Genus)) + 
  geom_bar(stat = "identity", position="fill")  +
  labs( x=" Genus", y="Percent Relative Abundance (%)") +  scale_fill_manual(name=NULL, values = c(many_col))  +  facet_wrap(~BiologicalMatter, scales= "free_x", nrow=1) +  
  theme_classic() + theme(strip.background = element_blank(), axis.text.x.bottom = element_text(angle = 180) +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold"), plot.title = element_text(size = 14, face = "bold",hjust = 0.5,),legend.text = element_text(size = 8), legend.box = "vertical") + theme(legend.text = element_text(size=12))+
                  theme(legend.key.size = unit(.9, 'cm')))  #title =expression(paste(italic(" Oribcella Spp."),"Microbial Community Composition Across Depth ")), + #
plot(testp)

```

# Hclutering
```{r}
## Use Distance matricies created above

# run hclust command
hc <- hclust(uni_flavo, method = "ward.D2")

# Get Tree tip label names 
tip_label <- flavo_df$sample_type

# Add tip labels to hclust object
hc$labels <- tip_label

## Create the Dendrogram
dend <- as.dendrogram(hc)

#dend <- color_branches(dend, k=9)


## clean up plot
dend %>% set("leaves_pch", 15) %>% 
  set("leaves_cex", 0.5) %>% # adjust the leave shape size
  set("labels_cex",1.0) %>% # adjust the leave label size
  color_branches(dend, k=4) %>%
  set("labels_col", k=4) %>%
  hang.dendrogram(hang_height = -0.002) %>% # hang level
  plot(main = "  Rhodobacteraceae \n   unifrac Dendrogram",
  horiz =  FALSE,  nodePar = list(cex = 0.01))


```
