---
title: "Global Coral Microbiome Project: Statistical Analysis and Modelling Workflow 2024"
authors: "Colin Howe, Sofia Roitman"
date: "01/17/2024"/
output: html_document
---

# First, Set working Directory load library and Import Data into Phyloseq
```{r warning=FALSE}
# Command for setting your work directory
setwd("C:/Users/Colin Howe/Desktop/PSU Doctoral Program/GCMP_Working/gcmp_qiime_input")

# Print your working directory

getwd()
#
```

## Load required libraries. 
```{r, message=FALSE, warning=FALSE} 
library(BiocManager)
library(phyloseq)
library(dplyr)
library(qiime2R)
library(DESeq2)
library(ComplexHeatmap)
library(ade4)
library(vegan)
library(RColorBrewer)
library(tidyverse)
library(ggplot2)
library(microbiome)
library(dendextend)
library(gplots)
library(ggrepel)
library(tidyr)
library(ggpubr)
library(cowplot)
library(RVAideMemoire)
library(ape)
library(microbiomeMarker)
```

# Import OTU Table Taxonomy and Tree from Qiime2 and import Metadata
```{r}

#Import from .qza file into a phyloseq object
asv <- qza_to_phyloseq(features = "gcmp_qiime_input/filtered_merged_table_v2.qza")

#### Import Metadata read.table
metadata <- read.table(file = "gcmp_qiime_input/gcmp_mapping30_2024.csv",header=T,comment.char="", row.names=1, sep=",")

### Import Tree file from biom output tree.nwk
tree2 <- read_tree("gcmp_qiime_input/tree.nwk")

### Import taxonomy from biom output as .tsv format using read.table
taxonomy <- read.table(file = "gcmp_qiime_input/taxonomy.tsv", sep = "\t", header = T ,row.names = 1)

# clean the taxonomy
##code referenced from Yan Hui: email me@yanh.org github: yanhui09
tax <- taxonomy %>%
  select(Taxon) %>% 
  separate(Taxon, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), "; ")

tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "k__",""),
                        Phylum = str_replace(tax[,2], "p__",""),
                        Class = str_replace(tax[,3], "c__",""),
                        Order = str_replace(tax[,4], "o__",""),
                        Family = str_replace(tax[,5], "f__",""),
                        Genus = str_replace(tax[,6], "g__",""),
                        Species = str_replace(tax[,7], "s__",""),
                        stringsAsFactors = FALSE)

tax.clean[is.na(tax.clean)] <- ""
tax.clean[tax.clean=="__"] <- ""

for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,7] != ""){
    tax.clean$Species[i] <- paste(tax.clean$Genus[i], tax.clean$Species[i], sep = " ")
  } else if (tax.clean[i,2] == ""){
    kingdom <- paste("Unclassified", tax.clean[i,1], sep = " ")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Unclassified", tax.clean[i,2], sep = " ")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Unclassified", tax.clean[i,3], sep = " ")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Unclassified", tax.clean[i,4], sep = " ")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Unclassified", tax.clean[i,5], sep = " ")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Unclassified ",tax.clean$Genus[i], sep = " ")
  }
}

# create matrix format for OTU and taxonomy table
OTU <- otu_table(as.matrix(asv), taxa_are_rows = TRUE)
tax1 = tax_table(as.matrix(tax.clean))

# Set metadata
SAMPLE <- sample_data(metadata)

# Read tree.nwk
TREE = read_tree(tree2)

# Create Working phyloseq object
main_coral <- phyloseq(OTU,tax1,SAMPLE,TREE)

library(gridExtra)
# Remove features that have only been classified at the kingdom level
phylo_noKing <-subset_taxa(main_coral, kingdom != "d__Eukaryota")

# Remove features that are identified as Chloroplast and Mitochondria or unclassified bacteria
phylo_noNA <- subset_taxa(phylo_noKing,Phylum != "NA", Order!= "Chloroplast" & Family != "Mitochondria" & Order !="Unclassified bacteroidia")

# Remove any samples with less than 1 read
phylo_coral = prune_samples(sample_sums(phylo_noNA)>1, phylo_noNA)
phylo_noNA
phylo_coral


```

### Subset Caribbean Phyloseq Objects for Comparative Analysis 
```{r}
## Create Phyloseq objects and subset dataset
#clear entire dataset
#rm(list=ls())

## Caribbean Dataset rarefication to 2500 per sample for downstream analysis
 # seed set to 111 to keep result reproductive

#### Subsample Robust clade
robust <-subset_samples(phylo_coral, ocean_area == "Caribbean" & complex_robust == "robust")
robust

# rarefy to even depth
robust.rarefied = rarefy_even_depth(robust, rngseed=111, sample.size=1000, replace=F)
robust.rarefied

## Subsample Robust Mucus
robust.m <-subset_samples(phylo_coral, ocean_area == "Caribbean" & complex_robust == "robust" & BiologicalMatter == "Coral Mucus")

# rarefy to even depth
robust.m.rarefied = rarefy_even_depth(robust.m, rngseed=111, sample.size=1000, replace=F)
robust.m.rarefied

## Subsample Robust Skeleton
robust.s <-subset_samples(phylo_coral, ocean_area == "Caribbean" & complex_robust == "robust"& BiologicalMatter == "Coral Skeleton")

# rarefy to even depth
robust.s.rarefied = rarefy_even_depth(robust.s, rngseed=111, sample.size=1000, replace=F)
robust.s.rarefied
robust.s

#### Subsample complex clade
complex <-subset_samples(phylo_coral, ocean_area == "Caribbean" & complex_robust == "complex")
complex

# rarefy to even depth
complex.rarefied = rarefy_even_depth(complex, rngseed=111, sample.size=1000, replace=F)
complex.rarefied

## Subsample complex mucus
complex.m <-subset_samples(phylo_coral, ocean_area == "Caribbean" & complex_robust == "complex" & BiologicalMatter == "Coral Mucus")
complex.m

# rarefy to even depth
complex.m.rarefied = rarefy_even_depth(complex.m, rngseed=111, sample.size=1000, replace=F)
complex.m.rarefied

## Subsample complex tissue # only 6 complex skeleton samples remain after filtering. So Tissue and Skeleton Samples were combined
complex.t <-subset_samples(phylo_coral, ocean_area == "Caribbean" & complex_robust == "complex"& BiologicalMatter == "Coral Tissue")
complex.t

# rarefy to even depth
complex.t.rarefied = rarefy_even_depth(complex.t, rngseed=111, sample.size=1000, replace=F)
complex.t.rarefied


#Sub_sample Corals Acropora
acro <-subset_samples(phylo_coral, ocean_area == "Caribbean" & family == "Acroporidae")
acro

# rarefy sample to event depth
acro.rarefied = rarefy_even_depth(acro, rngseed=111, sample.size=1000, replace=F)
acro.rarefied

## Subsample Meru * samples from both colombia and Panama may be an issue
meru <-subset_samples(phylo_coral, country == "Colombia" & family == "Merulinidae")
meru
# rarefy sample to event depth
meru.rarefied = rarefy_even_depth(meru, rngseed=111, sample.size=1000, replace=F)
meru.rarefied

## Subsample Meru
meru.s <-subset_samples(phylo_coral, ocean_area == "Caribbean" & family == "Merulinidae" & BiologicalMatter == "Coral Skeleton")
meru.s

# rarefy sample to event depth
meru.s.rarefied = rarefy_even_depth(meru.s, rngseed=111, sample.size=1000, replace=F)
meru.s.rarefied


#### Environmental Samples 
env_sample <- subset_samples(phylo_coral, env_sample == "yes" & ocean_area == "Caribbean")
env_sample

# keep result reproductive
env.rarefied = rarefy_even_depth(env_sample, rngseed=111, sample.size=1000, replace=F)
env.rarefied

## Plot distribution
glom_meru <- tax_glom(meru.rarefied, taxrank = 'Family', NArm = TRUE)## 21 samples
glom_meru_s <- tax_glom(meru.s.rarefied, taxrank = 'Family', NArm = TRUE)## 21 samples
glom_acro <- tax_glom(acro.rarefied, taxrank = 'Family', NArm = TRUE) ## 11 samples
glom_env <- tax_glom(env.rarefied, taxrank = 'Family', NArm = TRUE) ## 15 samples
glom_comp <- tax_glom(complex.rarefied, taxrank = 'Family', NArm = TRUE) ## 39 samples
glom_rob <- tax_glom(robust.rarefied, taxrank = 'Family', NArm = TRUE) ## 52 samples
glom_comp_m <- tax_glom(complex.m.rarefied, taxrank = 'Family', NArm = TRUE) ## 18 Samples
glom_comp_t <- tax_glom(complex.t.rarefied, taxrank = 'Family', NArm = TRUE) ## 6 Samples * Skeleton 
glom_rob_m <- tax_glom(robust.m.rarefied, taxrank = 'Family', NArm = TRUE) ## 17 Samples
glom_rob_s <- tax_glom(robust.s.rarefied, taxrank = 'Family', NArm = TRUE) ## 19 Samples 

```

### Subset Australia Phyloseq Objects for Comparative Analysis 
```{r}
## Create Phyloseq objects and subset dataset
#rm(list=ls())

## Australian Dataset rarefication to 1000 per sample for downstream analysis
 # seed set to 111 to keep result reproductive

#### Subsample Robust clade
robust <-subset_samples(phylo_coral,country_region == "Australia_Pacific" & complex_robust == "robust")
robust

# rarefy to even depth
robust.rarefied = rarefy_even_depth(robust, rngseed=111, sample.size=1000, replace=F)
robust.rarefied

## Subsample Robust Mucus
robust.m <-subset_samples(phylo_coral,country_region == "Australia_Pacific" & complex_robust == "robust" & BiologicalMatter == "Coral Mucus")

# rarefy to even depth
robust.m.rarefied = rarefy_even_depth(robust.m, rngseed=111, sample.size=1000, replace=F)
robust.m.rarefied

## Subsample Robust Skeleton
robust.s <-subset_samples(phylo_coral, country_region == "Australia_Pacific" & complex_robust == "robust"& BiologicalMatter == "Coral Skeleton")

# rarefy to even depth
robust.s.rarefied = rarefy_even_depth(robust.s, rngseed=111, sample.size=1000, replace=F)
robust.s.rarefied
robust.s

#### Subsample complex clade
complex <-subset_samples(phylo_coral, country_region == "Australia_Pacific" & complex_robust == "complex")
complex

# rarefy to even depth
complex.rarefied = rarefy_even_depth(complex, rngseed=111, sample.size=1000, replace=F)
complex.rarefied

## Subsample complex mucus
complex.m <-subset_samples(phylo_coral,  country_region == "Australia_Pacific" & complex_robust == "complex" & BiologicalMatter == "Coral Mucus")
complex.m

# rarefy to even depth
complex.m.rarefied = rarefy_even_depth(complex.m, rngseed=111, sample.size=1000, replace=F)
complex.m.rarefied

## Complex Coral Skeleton
complex.s <-subset_samples(phylo_coral, country_region == "Australia_Pacific" & complex_robust == "complex"& BiologicalMatter == "Coral Skeleton")
complex.s

# rarefy to even depth
complex.s.rarefied = rarefy_even_depth(complex.s, rngseed=111, sample.size=1000, replace=F)
complex.s.rarefied


#Sub_sample Corals Acropora
acro <-subset_samples(phylo_coral, country_region == "Australia_Pacific" & family == "Acroporidae")# & ocean == "Tasman Sea" & )
acro

# rarefy sample to event depth
acro.rarefied = rarefy_even_depth(acro, rngseed=111, sample.size=1000, replace=F)
acro.rarefied

## Subsample Meru
meru <-subset_samples(phylo_coral, country_region == "Australia_Pacific" & family == "Merulinidae")
meru
# rarefy sample to event depth
meru.rarefied = rarefy_even_depth(meru, rngseed=111, sample.size=1000, replace=F)
meru.rarefied

#### Environmental Samples 
env_sample <- subset_samples(phylo_coral, env_sample == "yes" &  country_region == "Australia_Pacific")
env_sample

# keep result reproductive
env.rarefied = rarefy_even_depth(env_sample, rngseed=111, sample.size=1000, replace=F)
env.rarefied

glom_meru_aus <- tax_glom(meru.rarefied, taxrank = 'Family', NArm = TRUE)## 91 samples
glom_acro_aus <- tax_glom(acro.rarefied, taxrank = 'Family', NArm = TRUE) ## 95 samples
glom_env_aus <- tax_glom(env.rarefied, taxrank = 'Family', NArm = TRUE) ## 10 samples
glom_comp_aus <- tax_glom(complex.rarefied, taxrank = 'Family', NArm = TRUE) ## 205 samples
glom_rob_aus <- tax_glom(robust.rarefied, taxrank = 'Family', NArm = TRUE) ## 247 samples
glom_comp_m_aus <- tax_glom(complex.m.rarefied, taxrank = 'Family', NArm = TRUE) ## 77 Samples
glom_comp_s_aus <- tax_glom(complex.s.rarefied, taxrank = 'Family', NArm = TRUE) ## 68 Samples 
glom_rob_m_aus <- tax_glom(robust.m.rarefied, taxrank = 'Family', NArm = TRUE) ## 99 Samples
glom_rob_s_aus <- tax_glom(robust.s.rarefied, taxrank = 'Family', NArm = TRUE) ## 77 Samples 

```

#Bruns et al. 2016 Neutral Model functions
```{r}
################define sncm.fit############
sncm.fit <- function(spp, pool=NULL, stats=TRUE, taxon=NULL){
	require(minpack.lm)
	require(Hmisc)
	require(stats4)
	
	options(warn=-1)

	#Calculate the number of individuals per community
	N <- mean(apply(spp, 1, sum))
	
	#Calculate the average relative abundance of each taxa across communities
	if(is.null(pool)){
		p.m <- apply(spp, 2, mean)
		p.m <- p.m[p.m != 0]
		p <- p.m/N
	} else {
		p.m <- apply(pool, 2, mean)
		p.m <- p.m[p.m != 0]
		p <- p.m/N
	}

#Calculate the occurrence frequency of each taxa across communities
spp.bi <- 1*(spp>0)
freq <- apply(spp.bi, 2, mean)
freq <- freq[freq != 0]
#Combine
C <- merge(p, freq, by=0)
C <- C[order(C[,2]),]
C <- as.data.frame(C)
C.0 <- C[!(apply(C, 1, function(y) any(y == 0))),] #Removes rows with any zero (absent in either source pool or local communities)
p <- C.0[,2]
freq <- C.0[,3]
names(p) <- C.0[,1]
names(freq) <- C.0[,1]

#Calculate the limit of detection
d = 1/N

##Fit model parameter m (or Nm) using Non-linear least squares (NLS)
m.fit <- nlsLM(freq ~ pbeta(d, N*m*p, N*m*(1-p), lower.tail=FALSE), start=list(m=0.1))
m.ci <- confint(m.fit, 'm', level=0.95)
	
##Fit neutral model parameter m (or Nm) using Maximum likelihood estimation (MLE)
sncm.LL <- function(m, sigma){
	R = freq - pbeta(d, N*m*p, N*m*(1-p), lower.tail=FALSE)
	R = dnorm(R, 0, sigma)
	-sum(log(R))
}
	
m.mle <- mle(sncm.LL, start=list(m=0.1, sigma=0.1), nobs=length(p))

##Calculate Akaike's Information Criterion (AIC)
aic.fit <- AIC(m.mle, k=2)
bic.fit <- BIC(m.mle)

##Calculate goodness-of-fit (R-squared and Root Mean Squared Error)
freq.pred <- pbeta(d, N*coef(m.fit)*p, N*coef(m.fit)*(1-p), lower.tail=FALSE)
Rsqr <- 1 - (sum((freq - freq.pred)^2))/(sum((freq - mean(freq))^2))
RMSE <- sqrt(sum((freq-freq.pred)^2)/(length(freq)-1))
	
pred.ci <- binconf(freq.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)
	
##Calculate AIC for binomial model
bino.LL <- function(mu, sigma){
R = freq - pbinom(d, N, p, lower.tail=FALSE)
R = dnorm(R, mu, sigma)
-sum(log(R))
	}
bino.mle <- mle(bino.LL, start=list(mu=0, sigma=0.1), nobs=length(p))
	
aic.bino <- AIC(bino.mle, k=2)
bic.bino <- BIC(bino.mle)
	
##Goodness of fit for binomial model
bino.pred <- pbinom(d, N, p, lower.tail=FALSE)
Rsqr.bino <- 1 - (sum((freq - bino.pred)^2))/(sum((freq - mean(freq))^2))
RMSE.bino <- sqrt(sum((freq - bino.pred)^2)/(length(freq) - 1))

bino.pred.ci <- binconf(bino.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)
	
##Calculate AIC for Poisson model
pois.LL <- function(mu, sigma){
	R = freq - ppois(d, N*p, lower.tail=FALSE)
	R = dnorm(R, mu, sigma)
	-sum(log(R))
}
pois.mle <- mle(pois.LL, start=list(mu=0, sigma=0.1), nobs=length(p))
	
aic.pois <- AIC(pois.mle, k=2)
bic.pois <- BIC(pois.mle)
	
##Goodness of fit for Poisson model
pois.pred <- ppois(d, N*p, lower.tail=FALSE)
Rsqr.pois <- 1 - (sum((freq - pois.pred)^2))/(sum((freq - mean(freq))^2))
RMSE.pois <- sqrt(sum((freq - pois.pred)^2)/(length(freq) - 1))

pois.pred.ci <- binconf(pois.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)

##Results
	if(stats==TRUE){
		fitstats <- data.frame(m=numeric(), m.ci=numeric(), m.mle=numeric(), maxLL=numeric(), binoLL=numeric(), poisLL=numeric(), Rsqr=numeric(), Rsqr.bino=numeric(), Rsqr.pois=numeric(), RMSE=numeric(), RMSE.bino=numeric(), RMSE.pois=numeric(), AIC=numeric(), BIC=numeric(), AIC.bino=numeric(), BIC.bino=numeric(), AIC.pois=numeric(), BIC.pois=numeric(), N=numeric(), Samples=numeric(), Richness=numeric(), Detect=numeric())
		fitstats[1,] <- c(coef(m.fit), coef(m.fit)-m.ci[1], m.mle@coef['m'], m.mle@details$value, bino.mle@details$value, pois.mle@details$value, Rsqr, Rsqr.bino, Rsqr.pois, RMSE, RMSE.bino, RMSE.pois, aic.fit, bic.fit, aic.bino, bic.bino, aic.pois, bic.pois, N, nrow(spp), length(p), d)
		return(fitstats)
	} else {
	A <- cbind(p, freq, freq.pred, pred.ci[,2:3], bino.pred, bino.pred.ci[,2:3])
	A <- as.data.frame(A)
		colnames(A) <- c('p', 'freq', 'freq.pred', 'pred.lwr', 'pred.upr', 'bino.pred', 'bino.lwr', 'bino.upr')
		if(is.null(taxon)){
			B <- A[order(A[,1]),]
		} else {
			B <- merge(A, taxon, by=0, all=TRUE)
			row.names(B) <- B[,1]
			B <- B[,-1]
			B <- B[order(B[,1]),]
		}
		return(B)
	}
}
##Identify above, below and neutral OTUs 

addf = function(freq, pred.upr, pred.lwr) {
  if(freq > pred.upr){
    model = "above"
  }
  else if(freq < pred.lwr){
    model = "below"
  }
  else if(pred.upr >= freq & freq >= pred.lwr){
    model = "neutral"
  }
  return(model)
}

```

## Neutral Model test
```{r}
###coral communities
set.seed(111)
coral <- prune_taxa(taxa_sums(robust) > 0, robust)
coral
otu_coral = as.data.frame(t(otu_table(coral)))

mod_w_coral = sncm.fit(otu_coral)
mod1_w_coral = sncm.fit(otu_coral,stats = F)

min(sample_sums(glom_comp_aus))

mod1_w_coral$model = mapply(addf, mod1_w_coral$freq, mod1_w_coral$pred.upr, mod1_w_coral$pred.lwr)  
length(which(mod1_w_coral$model == "above"))
length(which(mod1_w_coral$model == "below"))
length(which(mod1_w_coral$model == "neutral"))

mod1_w_coral$model <- factor(mod1_w_coral$model, 
                             levels = c("above", "neutral", "below"))

## Assign Taxonomy to table of significant microbes
coral_taxa = cbind(as(mod1_w_coral, "data.frame"), as(tax_table(glom_acro_aus)[rownames(mod1_w_coral), ], "matrix"))
#coral_taxa
write.csv(coral_taxa,"Australian Acroporidae Mucus Neutral Taxonomy table.csv",row.names = TRUE)


set.seed(111)
###seawater communities
#seawater <- subset_samples(physeq2, SampleType == "seawater")
seawater <- prune_taxa(taxa_sums(glom_meru_aus) >0, glom_meru_aus)
seawater

otu_seawater = as.data.frame(t(otu_table(seawater)))

mod_w_seawater = sncm.fit(otu_seawater)
mod1_w_seawater = sncm.fit(otu_seawater,stats = F)


mod1_w_seawater$model = mapply(addf, mod1_w_seawater$freq, mod1_w_seawater$pred.upr, mod1_w_seawater$pred.lwr)  
length(which(mod1_w_seawater$model == "above"))
length(which(mod1_w_seawater$model == "below"))
length(which(mod1_w_seawater$model == "neutral"))

mod1_w_seawater$model <- factor(mod1_w_seawater$model, 
                                levels = c("above", "neutral", "below"))

## Assign Taxonomy to table of significant microbes
env_taxa = cbind(as(mod1_w_seawater, "data.frame"), as(tax_table(glom_meru_aus)[rownames(mod1_w_seawater), ], "matrix"))

write.csv(env_taxa,"Australian Merulinidae Samples Neutral Taxonomy table.csv",row.names = TRUE)
```

## Plot Neutral Model
```{r}
#plot neuModel
p_neuM_coral=ggplot(mod1_w_coral, aes(x=log10(p), y=freq, color = model))+
  geom_point(shape=19, alpha=0.7, size=2.5)+
  geom_line(aes(x=log10(p), y=freq.pred),
            color="blue",linetype="solid",linewidth=1)+
  geom_line(aes(x=log10(p), y=pred.lwr),
            color="blue",linetype="dashed",linewidth=1)+
  geom_line(aes(x=log10(p), y=pred.upr),
            color="blue",linetype="dashed",linewidth=1)+
  scale_colour_manual(name="OTUs",
                      values=c("#E41A1C", "#999999", "#FF7F00"),
                      breaks=c("above", "neutral", "below"),
                      labels=c("Over-represented", "Neutrally-distributed", "Under-represented"))+
  xlab("Log10 (Mean Relative abundance)") + ylab("Occurance frequency")+
  theme_bw() +
  theme(axis.title.x = element_text(face="bold", size=15),
        axis.title.y = element_text(face="bold", size=15),
        axis.text.x = element_text(colour = "black", size=12),
        axis.text.y = element_text(colour = "black", size=12), 
        legend.title = element_text(size=11.5, face = "bold"),
        legend.text = element_text(size=11.5),
        legend.position="none")


lb1 <- mod_w_coral$Rsqr
lb2 <- mod_w_coral$m
lb3 <- expression(paste("Australian Acoporaidae Coral Samples"))

p_neuM_coral1 = p_neuM_coral + annotate("text", x = -3.5, y=0.7, size=6.5, label=lb1, 
                                        fontface="bold", color="black",
                                        parse=TRUE) +
  annotate("text", x = -3.5, y=0.6, size=6.5, label=lb2, 
           fontface="bold", color="black",
           parse=TRUE) +
  annotate("text", x = -3.5, y=0.9, size=6.5, label=lb3, 
           fontface="bold", color="black",
           parse=TRUE)


#plot neuModel
p_neuM_seawater=ggplot(mod1_w_seawater, aes(x=log10(p), y=freq, color = model))+
  geom_point(shape=19, alpha=0.7, size=2.5)+
  geom_line(aes(x=log10(p), y=freq.pred),
            color="blue",linetype="solid",size=1)+
  geom_line(aes(x=log10(p), y=pred.lwr),
            color="blue",linetype="dashed",size=1)+
  geom_line(aes(x=log10(p), y=pred.upr),
            color="blue",linetype="dashed",size=1)+
  scale_colour_manual(name="OTUs",
                      values=c("#E41A1C", "#999999", "#FF7F00"),
                      breaks=c("above", "neutral", "below"),
                      labels=c("Over-represented", "Neutrally-distributed", "Under-represented"))+
  xlab("Log10 (Mean Relative abundance)") + ylab("Occurance frequency")+
  theme_bw() +
  theme(axis.title.x = element_text(face="bold", size=15),
        axis.title.y = element_text(face="bold", size=15),
        axis.text.x = element_text(colour = "black", size=12),
        axis.text.y = element_text(colour = "black", size=12), 
        legend.title = element_text(size=11.5, face = "bold"),
        legend.text = element_text(size=11.5),
        legend.position="none")#http://127.0.0.1:24155/graphics/plot_zoom_png?width=1762&height=934

ev1 <- mod_w_seawater$Rsqr
ev2 <- mod_w_seawater$m
ev3 <- expression(paste("Australian\n Merulinidae Coral Samples "))

p_neuM_seawater1 = p_neuM_seawater + annotate("text", x = -3.5, y=0.7, size=6.5, label=ev1, 
                                              fontface="bold", color="black",
                                              parse=TRUE) +
annotate("text", x = -3.5, y=0.6, size=6.5, label=ev2, 
           fontface="bold", color="black",
           parse=TRUE) +
annotate("text", x = -3.5, y=0.9, size=5.5, label=ev3, 
           fontface="bold", color="black",
           parse=TRUE)

getwd()

mp_ncm = ggdraw() +
  draw_plot(p_neuM_coral1, x = 0.01, y = 0.1, width = 0.49, height = 0.8) + 
  draw_plot(p_neuM_seawater1, x = 0.51, y = 0.1, width = 0.49, height = 0.8) 
mp_ncm
#ggsave("neuModel.jpg", mp_ncm, width = 20, height = 10, dpi= 600)


```

# Caribbean DISTANCE MATRICES
```{r}
## Complex Corals 
bray_complex = phyloseq::distance(complex.rarefied, method="bray")
uni_complex = phyloseq::distance(complex.rarefied, method="unifrac")
wuni_complex = phyloseq::distance(complex.rarefied, method="wUnifrac")

# Create data frame for downstream analysis
comp_df <- data.frame(sample_data(complex.rarefied))

## Robust Corals 
bray_robust = phyloseq::distance(robust.rarefied, method="bray")
uni_robust = phyloseq::distance(robust.rarefied, method="unifrac")
wuni_robust = phyloseq::distance(robust.rarefied, method="wUnifrac")

# Create data frame for downstream analysis
rob_df <- data.frame(sample_data(robust.rarefied))


```

# Ordination from Distance Matrix and plot PCoA 
```{r}
##Calculate the PCoA on Complex Corals
rt.pcoa = ordinate(coral, method="PCoA", distance=wuni_coral)
# Set variables to zero for subsequent changes
#Ofra
pcoa<-0
pcoa <- plot_ordination(coral, rt.pcoa,  shape = "country_region",color="family")+  geom_point(size=3) + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + stat_ellipse(level = 0.95,type = "norm",aes(group=country_region))+ theme(legend.text=element_text(size=12)) + theme(axis.text=element_text(size=12, face = "bold"), axis.title=element_text(size=14,face="bold") + theme(text = element_text(size = 20,face = "bold"))) # title =expression(paste(italic(" Orbicella spp")," Bray-Curtis Distance PCoA across Depth")) scale_color_manual(values=c("#E31A1C","#1F78B4"))+
                      

```

## 1. PERMANOVAs Analysis
```{r}

#env material include mucus, tissue and skeleton, and outgroups without eDNA
distance_methods <-c("bray_coral","wuni_coral","uni_coral")
set.seed(129)

# This is what is called a for loop.        
for (i in distance_methods){ 
  form <- as.formula(paste(i,"country_region+family+BiologicalMatter+sequence_run", sep="~"))
  print(form)
 adonis2(form, data=coral_df)-> result
 print(result)
}
 

## Post-hoc comparison between Coral Compartments and Location
bray_coral_pairwise <-pairwise.perm.manova(bray_coral, coral_df$country_region,nperm=1000) 
uni_coral_pairwise <-pairwise.perm.manova(uni_coral,coral_df$biological_matter,nperm=1000)
wuni_coral_pairwise <-pairwise.perm.manova(wuni_coral,coral_df$biological_matter,nperm=1000)
uni_all_com_post <-pairwise.perm.manova(uni_all,all_out_df$political_area,nperm=1000)
uni_all_bio_post <-pairwise.perm.manova(uni_all,all_out_df$biological_matter,nperm=1000)

```

# Alpha Diversity BoxPlots Compartments & Family
```{r, boxplots}
###Once we have our sub_samples rarefied we can start with some diversity analysis  and visualization
### Alpha Diversity ** you can make multiple diversity box plots by switch x for sample variables

complex_diversity_compartment <- plot_richness(complex, x="family", color = "Genus", title = "Merulinidae Diversity Across compartment", measures = c("Shannon","Observed"),)+
  geom_boxplot() +
  theme_classic() +
  theme(strip.background = element_blank(), plot.title = element_text(hjust = 0.5), axis.text.x.bottom = element_text(angle = -90))
 
```

# BAR PLOTS: Coral Compartments
```{r, message=FALSE, warning=FALSE}
#### Use from agglomerated phyloseq objects from Neutral Model Section

# Transform sample counts to proportion out of 100
ps.all = transform_sample_counts(glom_comp, function(x) x/sum(x)*100)
ps.melt <- psmelt(ps.all)

# change to character for easy-adjusted level
ps.melt <- ps.melt %>%
  group_by(family,BiologicalMatter,host_genus,country_region,Family) %>%
  mutate(median=median(Abundance)) %>%
  arrange(desc(median))

# select group median > 1
keep <- unique(ps.melt$Order[ps.melt$median > 10])
ps.melt$Order[!(ps.melt$Order %in% keep)] <- "< %2 (Other)"

#to get the same rows together
#to get the same rows together
ps.melt_sum <- ps.melt %>%
  group_by(family,BiologicalMatter,Family) %>%
  summarize(Abundance = sum(Abundance), 
            mean = mean(Abundance), .groups = "drop") %>%
            mutate(Order= factor(Order),
                   Order= fct_reorder(Order,mean, .desc = TRUE),
                   Order= fct_shift(Order, n=1))
# Then Plot Results
# create an object with color blind safe orientation
many_col <- c("#A6CEE3", "#33A02C", "#B2DF8A",  "#1F78B4","#E31A1C","#6A3D9A", "#FF7F00","#FDBF6F", "#FB9A99", "#CAB2D6",  "#B15928","#1B9E77", "#D95F02", "#FFFF99","#E7298A", "#7570B3","#666666" ,"#66A61E", "#E6AB02","#A6761D", "#E41A1C", "#377EB8", "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF" ,"#D53E4F","#F46D43", "#FDAE61", "#FEE08B","#999999", "#E6F598","#ABDDA4","#3288BD","cyan1","#66C2A5")

testp <-0
testp <-ggplot(ps.melt, aes(x = family, y = Abundance, fill = Family)) + 
  geom_bar(stat = "identity", position="fill")  +
  labs( x=" family", y="Percent Relative Abundance (%)") +  scale_fill_manual(name=NULL, values = c(many_col))  +  facet_wrap(~BiologicalMatter, scales= "free_x", nrow=1) +  
  theme_classic() + theme(strip.background = element_blank(), axis.text.x.bottom = element_text(angle = 180) +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold"), plot.title = element_text(size = 14, face = "bold",hjust = 0.5,),legend.text = element_text(size = 8), legend.box = "vertical") + theme(legend.text = element_text(size=12))+
                  theme(legend.key.size = unit(.9, 'cm')))  #title =expression(paste(italic(" Oribcella Spp."),"Microbial Community Composition Across Depth ")), + #
plot(testp)

```

# Hclutering
```{r}
## Use Distance matricies created above

# run hclust command
hc_complex_muc <- hclust(uuni_acro_ske, method = "ward.D2")

# Get Tree tip label names 
coral_family <- acro_ske_df$country_region

# Add tip labels to hclust object
hc_complex_muc$labels <- coral_family

## Create the Dendrogram
dend <- as.dendrogram(hc_complex_muc)

#dend <- color_branches(dend, k=9)
plot(dend)

## clean up plot
dend %>% set("leaves_pch", 15) %>% 
  set("leaves_cex", 0.5) %>% # adjust the leave shape size
  set("labels_cex",1.0) %>% # adjust the leave label size
  color_branches(dend, k=4) %>%
  set("labels_col", k=4) %>%
  hang.dendrogram(hang_height = -0.002) %>% # hang level
  plot(main = "Caribbean Coral Microbiome Dendrogram",
  horiz =  TRUE,  nodePar = list(cex = 0.01))


```
