---
title: "Global Coral Microbiome Project: Statistical Analysis and Modelling Workflow 2024"
authors: "Colin Howe, Sofia Roitman"
start date: "01/17/2024"
current date: "03/29/2024"
output: html_document
---

# First, Set working Directory load library and Import Data into Phyloseq
```{r}
# Command for setting your work directory
knitr::opts_knit$set("C:/Users/Colin Howe/Desktop/PSU Doctoral Program/GCMP_Working/gcmp_analysis_2024/Analysis/Neutral_Model_Analysis/a1_neutrality_by_compartment_all")
                 
# Print your working directory

```

## Load required libraries. 
```{r, message=FALSE, warning=FALSE} 
library(BiocManager)
library(phyloseq)
library(dplyr)
library(qiime2R)
library(DESeq2)
library(ComplexHeatmap)
library(ade4)
library(vegan)
library(tidyverse)
library(ggplot2)
library(microbiome)
library(dendextend)
library(gplots)
library(tidyr)
library(RVAideMemoire)
library(ape)
library(microbiomeMarker)
library(paletteer)
library(RColorBrewer)
library(ggsignif)
library(ANCOMBC)
library(MiscMetabar)
```

# Import OTU Table Taxonomy and Tree from Qiime2 and import Metadata
```{r}
#Import from .qza file into a phyloseq object
asv <- qza_to_phyloseq(features = "./gcmp_qiime_input/merged_table.qza")

#### Import Metadata read.table
metadata <- read.table(file = "./gcmp_qiime_input/gcmp_complete_mapping2025.txt",header=T,comment.char="", row.names=1, sep="\t")

### Import Tree file from biom output tree.nwk
tree <- read_tree("gcmp_qiime_input/tree.nwk")

### Import taxonomy from biom output as .tsv format using read.table
taxonomy <- read.table(file = "gcmp_qiime_input/taxonomy.tsv", sep = "\t", header = T ,row.names = 1)

# clean the taxonomy
##code referenced from Yan Hui: email me@yanh.org github: yanhui09
tax <- taxonomy %>%
  select(Taxon) %>% 
  separate(Taxon, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), "; ")

tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "k__",""),
                        Phylum = str_replace(tax[,2], "p__",""),
                        Class = str_replace(tax[,3], "c__",""),
                        Order = str_replace(tax[,4], "o__",""),
                        Family = str_replace(tax[,5], "f__",""),
                        Genus = str_replace(tax[,6], "g__",""),
                        Species = str_replace(tax[,7], "s__",""),
                        stringsAsFactors = FALSE)

tax.clean[is.na(tax.clean)] <- ""
tax.clean[tax.clean=="__"] <- ""

for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,7] != ""){
    tax.clean$Species[i] <- paste(tax.clean$Genus[i], tax.clean$Species[i], sep = " ")
  } else if (tax.clean[i,2] == ""){
    kingdom <- paste("Unclassified", tax.clean[i,1], sep = " ")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Unclassified", tax.clean[i,2], sep = " ")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Unclassified", tax.clean[i,3], sep = " ")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Unclassified", tax.clean[i,4], sep = " ")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Unclassified", tax.clean[i,5], sep = " ")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Unclassified ",tax.clean$Genus[i], sep = " ")
  }
}

# create matrix format for OTU and taxonomy table
OTU <- otu_table(as.matrix(asv), taxa_are_rows = TRUE)
tax1 = tax_table(as.matrix(tax.clean))

# Set metadata
SAMPLE <- sample_data(metadata)

# Create Working phyloseq object
main_phylo <- phyloseq(OTU,tax1,SAMPLE,tree)

# visualize tax table
table(tax_table(main_phylo)[,"Kingdom"])

# remove unknown bacteria or unassigned
phylo_noking <-main_phylo %>%
  phyloseq::subset_taxa(!Kingdom %in% c("Unassigned","Unclassified d__Bacteria","d__Eukaryota")) %>%
  phyloseq::subset_taxa(!Phylum %in% c("Unclassified Unassigned","Unclassified d__Archaea","Unclassified d__Bacteria","Unclassified d__Eukaryota")) %>%
  phyloseq::subset_taxa(!Family %in% c("Mitochondria","Chloroplast"))

# verify present taxa in table
table(tax_table(phylo_noking)[,"Kingdom"])

# Remove any samples with less than 1 read
phylo = prune_samples(sample_sums(phylo_noking)>1, phylo_noking)
phylo

# Subset only corals from database
# Subset only corals from database
subject <-subset_samples(phylo, outgroup == "n" & tissue_compartment == "M")
subject

print(paste("Generating Rarefied Coral dataset..."))
# rarefy to even depth
rarefied = rarefy_even_depth(subject, rngseed=111, sample.size=1000, replace=F, trimOTUs = TRUE)
rarefied

###Agglomerate taxa to family 
glom <- tax_glom(rarefied, taxrank = 'Family', NArm = TRUE)
glom

```

#Bruns et al. 2016 Neutral Model functions
```{r}
################define sncm.fit############
sncm.fit <- function(spp, pool=NULL, stats=TRUE, taxon=NULL){
	require(minpack.lm)
	require(Hmisc)
	require(stats4)
	
	options(warn=-1)

	#Calculate the number of individuals per community
	N <- mean(apply(spp, 1, sum))
	
	#Calculate the average relative abundance of each taxa across communities
	if(is.null(pool)){
		p.m <- apply(spp, 2, mean)
		p.m <- p.m[p.m != 0]
		p <- p.m/N
	} else {
		p.m <- apply(pool, 2, mean)
		p.m <- p.m[p.m != 0]
		p <- p.m/N
	}

#Calculate the occurrence frequency of each taxa across communities
spp.bi <- 1*(spp>0)
freq <- apply(spp.bi, 2, mean)
freq <- freq[freq != 0]
#Combine
C <- merge(p, freq, by=0)
C <- C[order(C[,2]),]
C <- as.data.frame(C)
C.0 <- C[!(apply(C, 1, function(y) any(y == 0))),] #Removes rows with any zero (absent in either source pool or local communities)
p <- C.0[,2]
freq <- C.0[,3]
names(p) <- C.0[,1]
names(freq) <- C.0[,1]

#Calculate the limit of detection
d = 1/N

##Fit model parameter m (or Nm) using Non-linear least squares (NLS)
m.fit <- nlsLM(freq ~ pbeta(d, N*m*p, N*m*(1-p), lower.tail=FALSE), start=list(m=0.1))
m.ci <- confint(m.fit, 'm', level=0.95)
	
##Fit neutral model parameter m (or Nm) using Maximum likelihood estimation (MLE)
sncm.LL <- function(m, sigma){
	R = freq - pbeta(d, N*m*p, N*m*(1-p), lower.tail=FALSE)
	R = dnorm(R, 0, sigma)
	-sum(log(R))
}
	
m.mle <- mle(sncm.LL, start=list(m=0.1, sigma=0.1), nobs=length(p),"Nelder-Mead")

##Calculate Akaike's Information Criterion (AIC)
aic.fit <- AIC(m.mle, k=2)
bic.fit <- BIC(m.mle)

##Calculate goodness-of-fit (R-squared and Root Mean Squared Error)
freq.pred <- pbeta(d, N*coef(m.fit)*p, N*coef(m.fit)*(1-p), lower.tail=FALSE)
Rsqr <- 1 - (sum((freq - freq.pred)^2))/(sum((freq - mean(freq))^2))
RMSE <- sqrt(sum((freq-freq.pred)^2)/(length(freq)-1))
	
pred.ci <- binconf(freq.pred*nrow(spp), nrow(spp), alpha=0.01, method="wilson", return.df=TRUE)
	
##Calculate AIC for binomial model
bino.LL <- function(mu, sigma){
R = freq - pbinom(d, N, p, lower.tail=FALSE)
R = dnorm(R, mu, sigma)
-sum(log(R))
	}
bino.mle <- mle(bino.LL, start=list(mu=0, sigma=0.1), nobs=length(p))
	
aic.bino <- AIC(bino.mle, k=2)
bic.bino <- BIC(bino.mle)
	
##Goodness of fit for binomial model
bino.pred <- pbinom(d, N, p, lower.tail=FALSE)
Rsqr.bino <- 1 - (sum((freq - bino.pred)^2))/(sum((freq - mean(freq))^2))
RMSE.bino <- sqrt(sum((freq - bino.pred)^2)/(length(freq) - 1))

bino.pred.ci <- binconf(bino.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)
	
##Calculate AIC for Poisson model
pois.LL <- function(mu, sigma){
	R = freq - ppois(d, N*p, lower.tail=FALSE)
	R = dnorm(R, mu, sigma)
	-sum(log(R))
}
pois.mle <- mle(pois.LL, start=list(mu=0, sigma=0.1), nobs=length(p))
	
aic.pois <- AIC(pois.mle, k=2)
bic.pois <- BIC(pois.mle)
	
##Goodness of fit for Poisson model
pois.pred <- ppois(d, N*p, lower.tail=FALSE)
Rsqr.pois <- 1 - (sum((freq - pois.pred)^2))/(sum((freq - mean(freq))^2))
RMSE.pois <- sqrt(sum((freq - pois.pred)^2)/(length(freq) - 1))

pois.pred.ci <- binconf(pois.pred*nrow(spp), nrow(spp), alpha=0.05, method="wilson", return.df=TRUE)

##Results
	if(stats==TRUE){
		fitstats <- data.frame(m=numeric(), m.ci=numeric(), m.mle=numeric(), maxLL=numeric(), binoLL=numeric(), poisLL=numeric(), Rsqr=numeric(), Rsqr.bino=numeric(), Rsqr.pois=numeric(), RMSE=numeric(), RMSE.bino=numeric(), RMSE.pois=numeric(), AIC=numeric(), BIC=numeric(), AIC.bino=numeric(), BIC.bino=numeric(), AIC.pois=numeric(), BIC.pois=numeric(), N=numeric(), Samples=numeric(), Richness=numeric(), Detect=numeric())
		fitstats[1,] <- c(coef(m.fit), coef(m.fit)-m.ci[1], m.mle@coef['m'], m.mle@details$value, bino.mle@details$value, pois.mle@details$value, Rsqr, Rsqr.bino, Rsqr.pois, RMSE, RMSE.bino, RMSE.pois, aic.fit, bic.fit, aic.bino, bic.bino, aic.pois, bic.pois, N, nrow(spp), length(p), d)
		return(fitstats)
	} else {
	A <- cbind(p, freq, freq.pred, pred.ci[,2:3], bino.pred, bino.pred.ci[,2:3])
	A <- as.data.frame(A)
		colnames(A) <- c('p_abundance', 'freq', 'freq.pred', 'pred.lwr', 'pred.upr', 'bino.pred', 'bino.lwr', 'bino.upr')
		if(is.null(taxon)){
			B <- A[order(A[,1]),]
		} else {
			B <- merge(A, taxon, by=0, all=TRUE)
			row.names(B) <- B[,1]
			B <- B[,-1]
			B <- B[order(B[,1]),]
		}
		return(B)
	}
}
##Identify above, below and neutral OTUs 

addf = function(freq, pred.upr, pred.lwr, padj) {
  if(freq > pred.upr){
    model = "above"
  }
  else if(freq < pred.lwr){
    model = "below"
  }
  else if(pred.upr >= freq & freq >= pred.lwr & padj < 0.05){
    model = "neutral_significant"
  }
  else if(pred.upr >= freq & freq >= pred.lwr){
    model = "neutral"
  }
  return(model)
}
```

## Neutral Model test
```{r}
###coral communities
set.seed(111)
coral <- prune_taxa(taxa_sums(glom) > 0, glom)
coral
otu_coral = as.data.frame(t(otu_table(coral)))

## Neutral fit
mod_w_coral = sncm.fit(otu_coral)
mod1_w_coral = sncm.fit(otu_coral,stats = F)

min(sample_sums(glom))

mod1_w_coral %>% mutate(ci.mean = (pred.upr-pred.lwr)/2,
                        SE = (ci.mean/(1.96)),
                        zstat = (freq/SE),
                        p_value = 2*pnorm(-abs(zstat)),
                        #p_value = (exp(-0.717*zstat-0.416*zstat^2)),
                        padj = p.adjust(p_value, method = "bonferroni", n=length(mod1_w_coral$freq))) %>% as.data.frame() %>%
                        rownames_to_column("id")-> neutral_results_padj
                        #padj = (exp(-0.717*zstat-0.416*zstat^2))*length(mod1_w_coral$model)) -> neutral_results_padj

neutral_results_padj$model = mapply(addf, neutral_results_padj$freq, neutral_results_padj$pred.upr, neutral_results_padj$pred.lwr, neutral_results_padj$padj)  

length(which(neutral_results_padj$model == "above"))
length(which(neutral_results_padj$model == "below"))
length(which(neutral_results_padj$model == "neutral"))
length(which(neutral_results_padj$model == "neutral_significant"))

neutral_results_padj$model <- factor(neutral_results_padj$model, 
                             levels = c("above", "neutral", "below","neutral_significant"))

################# seawater ####################
set.seed(111)
###seawater communities
#seawater <- subset_samples(physeq2, SampleType == "seawater")
seawater <- prune_taxa(taxa_sums(glom_col) >0, glom_col)
seawater

otu_seawater = as.data.frame(t(otu_table(seawater)))

mod_w_seawater = sncm.fit(otu_seawater)
mod1_w_seawater = sncm.fit(otu_seawater,stats = F)


mod1_w_seawater$model = mapply(addf, mod1_w_seawater$freq, mod1_w_seawater$pred.upr, mod1_w_seawater$pred.lwr)  
length(which(mod1_w_seawater$model == "above"))
length(which(mod1_w_seawater$model == "below"))
length(which(mod1_w_seawater$model == "neutral"))

mod1_w_seawater$model <- factor(mod1_w_seawater$model, 
                                levels = c("above", "neutral", "below"))

## Assign Taxonomy to table of significant microbes
#env_taxa = cbind(as(mod1_w_seawater, "data.frame"), as(tax_table(glom_col12)[rownames(mod1_w_seawater), ], "matrix"))

#write.csv(env_taxa,"Colombia E12 Samples Neutral Taxonomy table.csv",row.names = TRUE)


```

## Plot Neutral Model
```{r}
#plot neuModel
p_neuM_coral=ggplot(neutral_results_padj, aes(x=log10(p_abundance), y=freq, color = model))+
  geom_point(shape=19, alpha=0.7, size=2.5)+
  geom_line(aes(x=log10(p_abundance), y=freq.pred),
            color="blue",linetype="solid",linewidth=1)+
  geom_line(aes(x=log10(p_abundance), y=pred.lwr),
            color="blue",linetype="dashed",linewidth=1)+
  geom_line(aes(x=log10(p_abundance), y=pred.upr),
            color="blue",linetype="dashed",linewidth=1)+
    scale_colour_manual(name="OTUs",
                      values=c("#E41A1C", "#999999", "#FF7F00","purple"),
                      breaks=c("above", "neutral", "below","neutral_significant"),
                      labels=c("Over-represented", "Neutrally-distributed", "Under-represented","Significant_low_weight"))+
  xlab("Log10 (Mean Relative abundance)") + ylab("Occurance frequency")+
  theme_bw() +
  theme(axis.title.x = element_text(face="bold", size=15),
        axis.title.y = element_text(face="bold", size=15),
        axis.text.x = element_text(colour = "black", size=12),
        axis.text.y = element_text(colour = "black", size=12), 
        legend.title = element_text(size=11.5, face = "bold"),
        legend.text = element_text(size=11.5),
        legend.position="none")

#expedition <-"E5"
biosample <-"Skeleton"
r <- round(mod_w_coral$Rsqr,3)
m_var <- round(mod_w_coral$m,3)
lb1 <- paste0("R_squared",":",r)
lb2 <- paste0("migration_rate (m):",m_var)
lb3 <- paste0(biosample,"_Neutral_Model")#expedition,"_",

p_neuM_coral1 = p_neuM_coral + annotate("text", x = -3.5, y=0.7, size=6.5, label=lb1, 
                                        fontface="bold", color="black",
                                        parse=TRUE) +
  annotate("text", x = -3.5, y=0.6, size=6.5, label=lb2, 
           fontface="bold", color="black",
           parse=TRUE) +
  annotate("text", x = -3.5, y=0.9, size=6.5, label=lb3, 
           fontface="bold", color="black",
           parse=FALSE)

#ggsave("Tissue_sig_neutral_model_plot.jpg", p_neuM_coral1, width = 20, height = 10, dpi= 600)

############ Seawater PLot #############
#plot neuModel
p_neuM_seawater=ggplot(mod1_w_seawater, aes(x=log10(p), y=freq, color = model))+
  geom_point(shape=19, alpha=0.7, size=2.5)+
  geom_line(aes(x=log10(p), y=freq.pred),
            color="blue",linetype="solid",size=1)+
  geom_line(aes(x=log10(p), y=pred.lwr),
            color="blue",linetype="dashed",size=1)+
  geom_line(aes(x=log10(p), y=pred.upr),
            color="blue",linetype="dashed",size=1)+
  scale_colour_manual(name="OTUs",
                      values=c("#E41A1C", "#999999", "#FF7F00"),
                      breaks=c("above", "neutral", "below"),
                      labels=c("Over-represented", "Neutrally-distributed", "Under-represented"))+
  xlab("Log10 (Mean Relative abundance)") + ylab("Occurance frequency")+
  theme_bw() +
  theme(axis.title.x = element_text(face="bold", size=15),
        axis.title.y = element_text(face="bold", size=15),
        axis.text.x = element_text(colour = "black", size=12),
        axis.text.y = element_text(colour = "black", size=12), 
        legend.title = element_text(size=11.5, face = "bold"),
        legend.text = element_text(size=11.5),
        legend.position="none")#http://127.0.0.1:24155/graphics/plot_zoom_png?width=1762&height=934

ev1 <- mod_w_seawater$Rsqr
ev2 <- mod_w_seawater$m
ev3 <- expression(paste("Colombia E2 & E12 \n Coral Samples"))

p_neuM_seawater1 = p_neuM_seawater + annotate("text", x = -3.5, y=0.7, size=6.5, label=ev1, 
                                              fontface="bold", color="black",
                                              parse=TRUE) +
annotate("text", x = -3.5, y=0.6, size=6.5, label=ev2, 
           fontface="bold", color="black",
           parse=TRUE) +
annotate("text", x = -3.5, y=0.9, size=5.5, label=ev3, 
           fontface="bold", color="black",
           parse=TRUE)

getwd()

mp_ncm = ggdraw() +
  draw_plot(p_neuM_coral1, x = 0.01, y = 0.1, width = 0.49, height = 0.8) + 
  draw_plot(p_neuM_seawater1, x = 0.51, y = 0.1, width = 0.49, height = 0.8) 
mp_ncm
ggsave(" Caribbean_Mucus Neutral Model Plot.jpg", p_neuM_coral1, width = 20, height = 10, dpi= 600)


```

## Psudo Tables 
```{r}
above_mucus = select(filter(nonneutral_tax, model == "above"), id,p_abundance,freq,padj,Phylum,Class,Order,Family,model) %>%
  mutate(compartment_neutrality = paste("mucus", model, sep = '_') )


## Subset glom OTU table
phyloseq::otu_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_table

## Subset glom tax table
phyloseq::tax_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_tax

## Subset glom tax table
phyloseq::sample_data(glom)%>%
        as.data.frame() -> glom_metadata

#subset otu table based on neutral results
above_mucus_table <-subset(glom_otu_table, id  %in%  above_mucus$id)
colnames(above_mucus_table) <- paste(colnames(above_mucus_table), "above", sep = "_")
above_mucus_tax 

```
## Neutral Phyloseq Object      
```{r}


## ancombc analysis
library(ANCOMBC)


## Merging psudo_sample tables taxonomy and mapping files across compartments
#Mucus
mucus_table <-read.csv("./Mucus/Mucus_combinded_table.csv", row.names=1, check.names=FALSE)
mucus_tax <-read.csv("./Mucus/Mucus_qiime_neutral_taxonomy.csv", row.names=1, check.names=FALSE)
mucus_mapping <-read.csv("./Mucus/Mucus_qiime_neutral_mapping.csv", row.names=1, check.names=FALSE)

#Tissue
Tissue_table <-read.csv("./Tissue/Tissue_combinded_table.csv", row.names=1, check.names=FALSE)
Tissue_tax <-read.csv("./Tissue/Tissue_qiime_neutral_taxonomy.csv", row.names=1, check.names=FALSE)
Tissue_mapping <-read.csv("./Tissue/Tissue_qiime_neutral_mapping.csv", row.names=1, check.names=FALSE)

#Skeleton
Skeleton_table <-read.csv("./Skeleton/Skeleton_combinded_table.csv", row.names=1, check.names=FALSE)
Skeleton_tax <-read.csv("./Skeleton/Skeleton_qiime_neutral_taxonomy.csv", row.names=1, check.names=FALSE)
Skeleton_mapping <-read.csv("./Skeleton/Skeleton_qiime_neutral_mapping.csv", row.names=1, check.names=FALSE)

## format import files 
mucus_table[is.na(mucus_table)] <-0

## Filter row artifacts
m_map <-mucus_mapping[-c(1,2,3),]
t_map <-Tissue_mapping[-c(1,2,3),]
s_map <-Skeleton_mapping[-c(1,2,3),]

# create matrix format for OTU and taxonomy table
MOTU <- otu_table(as.matrix(mucus_table), taxa_are_rows = TRUE)
Mtax1 = tax_table(as.matrix(mucus_tax))



## Filter row artifacts
m_map <-mucus_mapping[-c(1,2,3),]
t_map <-Tissue_mapping[-c(1,2,3),]
s_map <-Skeleton_mapping[-c(1,2,3),]

# Create new column with compartment and neutrality
m_map %>% mutate(compartment_neutrality  = paste(compartment, model, sep = '_')) -> skeleton_map
t_map %>% mutate(compartment_neutrality  = paste(compartment, model, sep = '_')) -> tissue_map
s_map %>% mutate(compartment_neutrality  = paste(compartment, model, sep = '_')) -> mucus_map



# Set metadata
MSAMPLE <- sample_data(mucus_map)

# Create Working phyloseq object
mucus_phylo <- phyloseq(MOTU,Mtax1,MSAMPLE,tree)
mucus_phylo
clean_mucus <- prune_taxa(taxa_sums(mucus_phylo)>1,mucus_phylo)


## Merge OTU Tables across compartments
merge(mucus_table,Tissue_table, by=0, all=T)%>% 
  merge(Skeleton_table) -> test_table 

test_table[is.na(test_table)] <-0

colnames(merged_tax)[1] <- ""

## Merge Taxonomy files 
merge(mucus_tax,Tissue_tax, by=0, all=T) %>% 
  merge(Skeleton_tax) -> test_tax 
colnames(test_tax)
## coalesce taxonomy columnss
merged_tax <- test_tax %>% mutate(
                              Taxon = coalesce(Taxon.x,Taxon.y),
                              Consensus = coalesce(Consensus.x,Consensus.y)) %>%
                              select(Row.names,Taxon,Consensus)
colnames(merged_tax)[1] <- ""

## Merge each mapping file
merge(mucus_map,tissue_map, by=0, all=T) %>% 
  merge(skeleton_map) -> merge_map 

## Coalesce colummns
merge_map %>% mutate(compartment = coalesce(compartment.x,compartment.y),
                     model = coalesce(model.x,model.y),
                     compartment_neutrality = coalesce(compartment_neutrality.x, compartment_neutrality.y)) %>% 
                    select(Row.names,model,compartment,compartment_neutrality)-> combined_map

```

## Linear regression analysis
```{r}
knitr::opts_knit$set(root.dir = 'C:/Users/Colin Howe/Desktop/PSU Doctoral Program/GCMP_Working/gcmp_analysis_2024/Neutral Model Analysis')
remove(results_table)
results_table <- read.table(file = "gcmp_qiime_input/neutral_results_table.csv",header=T,comment.char="",sep=",")

compartment = results_table$Compartment
as.character(compartment)
ocean = results_table$Ocean
region = results_table$Region
Rsquared = results_table$R2
faith = results_table$Faith.PD
faithsr = results_table$Faiths.SR
n_samples = results_table$n_samples
region = factor(results_table$Region)
below = results_table$n_belowMean
above = results_table$n_aboveMean
expedition = results_table$expedition
migration = results_table$m

category_migration_test <- lm(migration ~ compartment+region+faith+faithsr+ocean)
summary(category_migration_test)

category_Rquared_test <- lm(Rsquared ~ compartment+region+faith+faithsr+ocean)
summary(category_Rquared_test) 


## Plot Linear regression
faith_m_coe <- paste0("fatihpd_pvalue:\n",round(category_migration_test$coefficients[7],7))

lm_plot = plot(faith,migration, pch = 16, cex=1.3, col="#009197", main = "Migration and Coral Faith Pd", xlab = "Coral Species Diversity (faith Pd)", ylab ="migration (m)") + text(x=1500, y=.07, label = faith_m_coe) + abline(lm(migration ~ faith), col="#EC8C5C") 

#write.table(migration_regression_results,file="migration_regression_results.txt", sep = "\t",row.names = TRUE)
#write.table(Rsquared_regression_results,file="R2_regression_results.txt", sep = "\t",row.names = TRUE)

##  Compartment Boxplot with R2 
coral_color =c("#009197","#EC8C5C","#8E3B97")

testp <- ggplot(results_table, aes(y=Rsquared, x=compartment)) + geom_boxplot(aes(color=compartment)) + theme_classic() + geom_point(aes(color=compartment)) +labs(xlab="R2", ylab="Coral Compartments", title ="Compartment Model Fit") + theme(plot.title = element_text(hjust = 0.5)) + scale_color_manual(values=c("#009197","#EC8C5C","#8E3B97")) + geom_signif(comparisons = list(c("Mucus", "Tissue"), c("Mucus", "Skeleton"),c("Tissue", "Skeleton")), y_position = c(.85,1, .75))

```

## Faith Pd  Analysis
```{r}
coral_object <-subset_samples(coral, BiologicalMatter == "Coral Mucus") # all corals without out groups, environmental samples & subset expeditions (i.e, singapore).
coral_object

# pull out sample data 
phyloseq::sample_data(coral_object) %>%  
    group_by(sample_name_backup, expedition_number,BiologicalMatter) %>%
  as.data.frame() %>%
  select(sample_name_backup, expedition_number,BiologicalMatter) -> microbial_faith

#Microbial Phyloseq Pd analysis 
estimate_pd(coral_object) %>%
  as.data.frame() -> microbial_faith_pd

microbial_faith$faith_pd <- microbial_faith_pd$PD
microbial_faith$faith_SR <- microbial_faith_pd$SR

# Create data frame from sample data 
phyloseq::sample_data(coral_object) %>%  
  group_by(expedition_number, BiologicalMatter,Huang_Roy_tree_name) %>%
  as.data.frame() %>%
  select(expedition_number, BiologicalMatter,Huang_Roy_tree_name)-> test_df

# Create a new column titled eco to join expedition and biological matter
test_df$eco <-paste(test_df$expedition_number,test_df$BiologicalMatter, sep = "_")

#group and count total for each unique group maintaining NA values
test_df %>% group_by(eco, Huang_Roy_tree_name) %>%  summarise(counts=n()) %>%
  ungroup %>%
  complete(nesting(eco),
           nesting(Huang_Roy_tree_name),
           fill = list(quantity = 0)) -> test_table

# Fill NA with 0
test_table[is.na(test_table)] <-0

#Build Matrix
e <- unique(test_table$eco) 
t <- unique(test_table$Huang_Roy_tree_name) 
c <- test_table$counts

test_matrix <- matrix(c, nrow = length(e), ncol = length(t), byrow=TRUE)
rownames(test_matrix) = e
colnames(test_matrix) = t

# clean data set to match each other
clean_tree <- match.phylo.comm(phy = coral_tree, comm = test_matrix)$phy
clean_comm <- match.phylo.comm(phy = coral_tree, comm = test_matrix)$comm

coral_faith_pd <- pd(clean_comm, clean_tree, include.root=TRUE)
coral_faithpd_reorded <-coral_faith_pd[order(coral_faith_pd$PD, decreasing=TRUE),] 

write.table(microbial_faith, file ="./carib_mucus_microbial_faith_pd_table.csv", sep = ",",col.names = TRUE)

write.csv(coral_faithpd_reorded, file ="./carib_mucus_Host_faith_pd_table.csv" ,row.names = TRUE)
```

# BAR PLOTS: Coral Compartments
```{r}

##import neutral Model results tables

tissue_table <- read.table(file ="gcmp_qiime_input/Tissue_neutral_model.csv",header=T,comment.char="", row.names=1, sep=",")
mucus_table <- read.table(file ="gcmp_qiime_input/Mucus_neutral_model.csv",header=T,comment.char="", row.names=1, sep=",")
skeleton_table <- read.table(file ="gcmp_qiime_input/Skeleton_neutral_model.csv",header=T,comment.char="", row.names=1, sep=",")
combined_table <- read.table(file ="gcmp_qiime_input/combined_neutral_table.csv",header=T,comment.char="", row.names=1, sep=",")

sig_tissue <-filter(tissue_table, padj <=0.05)
sig_skeleton <-filter(skeleton_table, padj <=0.05)
sig_mucus <-filter(mucus_table, padj <=0.05)
sig_combined <-filter(combined_table, padj <=0.05)
filter(sig_combined, Phylum == "Proteobacteria") %>%
            mutate(Order= factor(Order),
                   Order= fct_reorder(Order,p, .desc = FALSE)) -> protobacteria



many_col <- c("#A6CEE3", "#33A02C","#E7298A","#FDBF6F", "#B2DF8A", "#1F78B4","#E31A1C","#6A3D9A", "#FF7F00","#1B9E77","#B15928", "#FB9A99", "#CAB2D6", "#377EB8","#D95F02", "#7570B3","#FFFF99","#666666" ,"#66A61E", "#E6AB02","#A6761D","black", "#CAB2D6",  "#B15928","#FFFF99", "#D95F02", "#E7298A", "#7570B3","#666666" , "#E6AB02", "brown3","#A6761D", "#377EB8", "#4DAF4A" ,"#984EA3", "#FFFF33","#FF7F00","cyan3", "#4DAF4A" ,"#FF7F00","#984EA3", "#A65628","#1B9E77", "#F781BF", "#F46D43","#ABDDA4", "#FDAE61","black", "#FEE08B","#E6F598","#3288BD","deeppink","#66C2A5","cyan1","red3", "#E69F00","#56B4E9","#999999","#F0E442","#0072B2","#D55E00","#D53E4F","yellow","skyblue","coral4","#A6CEE3","white","darkkhaki","brown1","#FFFF33","chocolate","darkorchid2","#FF7F00","#1B9E77","#66A61E","#FDBF6F", "#FB9A99","#A6CEE3", "#33A02C","#E7298A","#FDBF6F", "#B2DF8A", "#1F78B4","#E31A1C","#6A3D9A", "#FF7F00","#1B9E77","#B15928", "#FB9A99", "#CAB2D6", "#377EB8","#D95F02", "#7570B3","#FFFF99","#666666" ,"#66A61E", "#E6AB02")


proto_bar <- ggplot(data = protobacteria, aes(x = model, y = p, fill = Order)) +
         geom_bar(stat = "identity", position="stack") + 
  facet_wrap(~sample_type, scales= "free_x", nrow=1) +
  labs( x=" family", y="average relative abundance (p)") + guides(fill = guide_legend(title = "Order"))+ ggtitle("Non-Neutral Proteobacteria Bargraph") + scale_fill_manual(name=NULL, values = c(many_col)) 


all_bar <- ggplot(data = sig_combined, aes(x = model, y = p, fill = Phylum)) +
         geom_bar(stat = "identity", position="stack") + 
  facet_wrap(~sample_type, scales= "free_x", nrow=1) +
  labs( x=" family", y="average relative abundance (p)") + ggtitle("Compartment Neutral Bargraph") + scale_fill_manual(name=NULL, values = c(many_col)) 

mucus_bar <- ggplot(data = sig_mucus, aes(x = model, y = p, fill = Phylum)) +
         geom_bar(stat = "identity", position="stack") + 
  labs( x=" family", y="average relative abundance (p)") + ggtitle("Mucus Neutral Bargraph") + scale_fill_manual(name=NULL, values = c(many_col)) 

tissue_bar <- ggplot(data = sig_tissue, aes(x = model, y = p, fill = Phylum)) +
         geom_bar(stat = "identity", position="stack")  +
  labs( x=" family", y="average relative abundance (p)") + ggtitle("Tissue Neutral Bargraph") +  scale_fill_manual(name=NULL, values = c(many_col)) 

skeleton_bar <- ggplot(data = sig_skeleton, aes(x = model, y = p, fill = Phylum)) +
         geom_bar(stat = "identity", position="stack")  +
  labs( x=" family", y="average relative abundance (p)") + ggtitle("Skeleton Neutral Bargraph") +  scale_fill_manual(name=NULL, values = c(many_col)) 

```

# Calculate DISTANCE MATRICES
```{r}
## Mucus

bray_muc = phyloseq::distance(clean_mucus, method="bray")
uni_muc = phyloseq::distance(clean_mucus, method="unifrac")
wuni_muc= phyloseq::distance(clean_mucus, method="wUnifrac")

# Create data frame for downstream analysis
muc_df <- data.frame(sample_data(clean_mucus))


```

# Ordination from Distance Matrix and PCoA Plots 
```{r}
##Calculate the PCoA on Complex Corals
rt.pcoa = ordinate(mucus_phylo, method="PCoA", distance=uni_muc)
# Set variables to zero for subsequent changes
#Ofra
pcoa<-0
pcoa <- plot_ordination(mucus_phylo, rt.pcoa ,color="model")+  geom_point(size=3) + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + stat_ellipse(level = 0.95,type = "norm",aes(group=model))+ theme(legend.text=element_text(size=12)) + theme(axis.text=element_text(size=12, face = "bold"), axis.title=element_text(size=14,face="bold") + theme(text = element_text(size = 20,face = "bold"))) # title =expression(paste(italic(" Orbicella spp")," Bray-Curtis Distance PCoA across Depth")) scale_color_manual(values=c("#E31A1C","#1F78B4"))+
                      

```

# Alpha Diversity BoxPlots Compartments & Family
```{r, boxplots}
###Once we have our sub_samples rarefied we can start with some diversity analysis  and visualization
### Alpha Diversity ** you can make multiple diversity box plots by switch x for sample variables

complex_diversity_compartment <- plot_richness(complex, x="family", color = "Genus", title = "Merulinidae Diversity Across compartment", measures = c("Shannon","Observed"),)+
  geom_boxplot() +
  theme_classic() +
  theme(strip.background = element_blank(), plot.title = element_text(hjust = 0.5), axis.text.x.bottom = element_text(angle = -90))
 
```

## 1. PERMANOVAs Analysis
```{r}

#env material include mucus, tissue and skeleton, and outgroups without eDNA
distance_methods <-c("bray_muc","wuni_muc","uni_muc")
set.seed(129)

# This is what is called a for loop.        
for (i in distance_methods){ 
  form <- as.formula(paste(i,"model", sep="~"))
  print(form)
 adonis2(form, data=muc_df)-> result
 print(result)
}
 

## Post-hoc comparison between Coral Compartments and Location
bray_coral_pairwise <-pairwise.perm.manova(bray_coral, coral_df$country_region,nperm=1000) 
uni_coral_pairwise <-pairwise.perm.manova(uni_coral,coral_df$biological_matter,nperm=1000)
wuni_coral_pairwise <-pairwise.perm.manova(wuni_coral,coral_df$biological_matter,nperm=1000)
uni_all_com_post <-pairwise.perm.manova(uni_all,all_out_df$political_area,nperm=1000)
uni_all_bio_post <-pairwise.perm.manova(uni_all,all_out_df$biological_matter,nperm=1000)

```

# Hclutering
```{r}
## Use Distance matricies created above

# run hclust command
hc_complex_muc <- hclust(wuni_flavo, method = "ward.D2")

# Get Tree tip label names 
coral_family <- flavo_dd$BiologicalMatter

# Add tip labels to hclust object
hc_complex_muc$labels <- coral_family

## Create the Dendrogram
dend <- as.dendrogram(hc_complex_muc)

#dend <- color_branches(dend, k=9)


## clean up plot
dend %>% set("leaves_pch", 15) %>% 
  set("leaves_cex", 0.5) %>% # adjust the leave shape size
  set("labels_cex",1.0) %>% # adjust the leave label size
  color_branches(dend, k=4) %>%
  set("labels_col", k=4) %>%
  hang.dendrogram(hang_height = -0.002) %>% # hang level
  plot(main = "Caribbean Actinomarinaceae Dendrogram",
  horiz =  FALSE,  nodePar = list(cex = 0.01))


```
## Parking lot
```{r}
## Subset glom OTU table
phyloseq::otu_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_table

## Subset glom tax table
phyloseq::tax_table(glom)%>%
        as.data.frame()%>%
        rownames_to_column("id") -> glom_otu_tax

## Subset glom tax table
phyloseq::sample_data(glom)%>%
        as.data.frame() -> glom_metadata

# Connnect taxonomy information to neutral  model results
inner_join(neutral_results_padj,glom_tax, by = "id") %>% as.data.frame() -> nonneutral_tax


#subset otu table based on neutral results
neutral_table <-subset(glom_otu_table, id  %in%  neutral_fam_df$id)

#subset taxonomy table based on neutral results
neutral_tax <- subset(glom_otu_tax, id  %in%  neutral_fam_df$id)# & Family %in% neutral_fam_df$Family)

#subset Metadata based on neutral results
neutral_sample_name <- colnames(neutral_table)
neutral_meta <-subset(glom_metadata,sample_name_backup %in% neutral_sample_name)
neutral_metadata <- sample_data(neutral_meta)

## This approch helps transpose the data without adding numbers as colummn id
test_table <- setNames(data.frame(t(glom_otu_table[,-1])),glom_otu_table[,1])

# create matrix format for OTU and taxonomy table
OTU2 <- otu_table(data.matrix(test_table, rownames.force = NA), taxa_are_rows = TRUE)
tax2 = tax_table(as.matrix(neutral_tax))

# Create Working phyloseq object
neutral_phylo <- phyloseq(OTU2,tax2,neutral_metadata)

neutral_phylo



```